---
title: "Colombano AWS model checks"
author: "Denise Colombano"
date: "October 29, 2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Load libraries
```{r setup, include=FALSE}
install.packages("devtools")
devtools::install_github("thomasp85/patchwork")

library(tidyverse)
library(lubridate)
library(patchwork)

library(brms)
library(broom)
library(bayesplot)
library(modelr)
library(loo)

library(rstan)
library(StanHeaders)
library(parallel)

rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores(), loo.cores = parallel::detectCores())
getOption("mc.cores", "loo.cores")
```


# re-import data
Import twilight, in-channel flows, and species-specific dataframes in the script, "Colombano AWS modeling". Splittail= dst, Striped bass= dsb, Tule perch= dtp.


## Splittail
```{r}
model00st <- readRDS("Models/Splittail/brms_zip_00st_subset.rds")
model0st <- readRDS("Models/Splittail/brms_zip_0st_subset.rds")
model1st <- readRDS("Models/Splittail/brms_zip_1st_subset.rds")
model2st <- readRDS("Models/Splittail/brms_zip_2st_subset.rds")
model3st <- readRDS("Models/Splittail/brms_zip_3st_subset.rds")
model4st <- readRDS("Models/Splittail/brms_zip_4st_subset.rds")
model5st <- readRDS("Models/Splittail/brms_zip_5st_subset.rds")
model6st <- readRDS("Models/Splittail/brms_zip_6st_subset.rds")
model7st <- readRDS("Models/Splittail/brms_zip_7st_subset.rds")
model8st <- readRDS("Models/Splittail/brms_zip_8st_subset.rds")
model9st <- readRDS("Models/Splittail/brms_zip_9st_subset.rds")
model10st <- readRDS("Models/Splittail/brms_zip_10st_subset.rds")
model11st <- readRDS("Models/Splittail/brms_zip_11st_subset.rds")
model12st <- readRDS("Models/Splittail/brms_zip_12st_subset.rds")
model13st <- readRDS("Models/Splittail/brms_zip_13st_subset.rds")
model14st <- readRDS("Models/Splittail/brms_zip_14st_subset.rds")
model15st <- readRDS("Models/Splittail/brms_zip_15st_subset.rds")
model16st <- readRDS("Models/Splittail/brms_zip_16st_subset.rds")
model17st <- readRDS("Models/Splittail/brms_zip_17st_subset.rds")
model18st <- readRDS("Models/Splittail/brms_zip_18st_subset.rds")
```

```{r}
# check for problematic observations: pareto k estimates (k > 0.7)

LOO(model00st) # OK
LOO(model0st) # OK
LOO(model1st) # OK
LOO(model2st) # OK
LOO(model3st) # OK
LOO(model4st) # OK
LOO(model5st) # OK
LOO(model6st) # OK
LOO(model7st) # OK
LOO(model8st) # OK
LOO(model9st) # OK
LOO(model10st) # OK
LOO(model11st) # OK
LOO(model12st) # OK
LOO(model13st) # OK
LOO(model14st) # OK
LOO(model15st) # OK
LOO(model16st) # OK
LOO(model17st) # OK
LOO(model18st) # OK

# no need to reloo splittail models
```

Add loo criterion
```{r}
model000st <- add_loo(model000st)
model00st <- add_loo(model00st)
model0st <- add_loo(model0st)
model1st <- add_loo(model1st)
model2st <- add_loo(model2st)
model3st <- add_loo(model3st)
model4st <- add_loo(model4st)
model5st <- add_loo(model5st)
model6st <- add_loo(model6st)
model7st <- add_loo(model7st)
model8st <- add_loo(model8st)
model9st <- add_loo(model9st)
model10st <- add_loo(model10st)
model11st <- add_loo(model11st)
model12st <- add_loo(model12st)
model13st <- add_loo(model13st)
model14st <- add_loo(model14st)
model15st <- add_loo(model15st)
model16st <- add_loo(model16st)
model17st <- add_loo(model17st)
model18st <- add_loo(model18st)
model19st <- add_loo(model19st)
```


### model results
https://bookdown.org/connect/#/apps/1850/access
Solomon Kurz code: Rethinking 6.5.1 and 6.5.2
```{r}
# brms compare loo
st_loo <- loo_compare(model00st, model0st, model1st, model2st, model3st, model4st, model5st, model6st, model7st, model8st, model9st, model10st, model11st, model12st, model13st, model14st, model15st, model16st, model17st, model18st, model19st, model20st, model21st)
st_loo

st_loo_df <- print(st_loo, simplify=F, digits=1)
st_loo_df

st_loo_df2 <- st_loo_df %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(model=rowname)
View(st_loo_df2)

st_loo2 <- cbind(loo_diff = st_loo[, 1] * -2,
      se        = st_loo[, 2] * 2) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(model=rowname)
View(st_loo2)

st_loo2 %>% write_csv("Data_output/SPT_looic.csv")

# model weights
# regular weights
st_mw_loo <- model_weights(model00st, model0st, model1st, model2st, model3st, model4st, model5st, model6st, model7st, model8st, model9st, model10st, model11st, model12st, model13st, model14st, model15st, model16st, model17st, model18st, model19st, model20st, model21st, weights = "loo")
View(st_mw_loo) 

st_mw_loo <- data.frame(st_mw_loo) %>% rownames_to_column() %>% 
  rename(model=rowname, model_weights=st_mw_loo) %>% 
  mutate(model_weights=round(model_weights, digits=2)) %>% 
  arrange(desc(model_weights))
st_mw_loo
View(st_mw_loo)

st_mw_loo %>% write_csv("Data_output/SPT_Model_weights_loo.csv")

# stacked weights
st_mw_loo2 <- model_weights(model00st, model0st, model1st, model2st, model3st, model4st, model5st, model6st, model7st, model8st, model9st, model10st, model11st, model12st, model13st, model14st, model15st, model16st, model17st, model18st, model19st, model20st, model21st, weights = "loo2") # loo2 for Bayesian stacked weights
View(st_mw_loo2) 

st_mw_loo2 <- data.frame(st_mw_loo2) %>% rownames_to_column() %>% 
  rename(model=rowname, stacked_weights=st_mw_loo2) %>% 
  mutate(stacked_weights=round(stacked_weights, digits=2)) %>% 
  arrange(desc(stacked_weights))
st_mw_loo2
View(st_mw_loo2)

st_mw_loo2 %>% write_csv("Data_output/SPT_Stacked_weights_loo2.csv")

# compare regular vs. stacked weights
st_tbl_loo <- st_loo_df2 %>% #looic
  left_join(st_mw_loo, by="model") %>% # regular model weights
  left_join(st_mw_loo2, by="model") # stacked model weights
View(st_tbl_loo)

st_tbl_loo %>% write_csv("Data_output/SPT_Model_looic_compare_loo_vs_loo2.csv")
```

### effect plots
```{r}
library(brms)

# marginal effects and splines plots
# model 15
mest15 <- marginal_effects(model15st, effects="ills", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mest15)

msst15 <- marginal_smooths(model15st, resolution=1000)
msst15b <- marginal_smooths(model15st, resolution=1000, probs=c(0.1, 0.9))
msst15c <- marginal_smooths(model15st, resolution=1000, probs=c(0.25, 0.75))

disk_interval_st <- msst15b$"mu: s(ills,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

disk_interval_st2 <- msst15c$"mu: s(ills,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

st_plot15 <- plot(msst15)[[1]]+
  geom_ribbon(data=disk_interval_st, fill="darkgray")+
  geom_ribbon(data=disk_interval_st2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="% Disk illumination", y="mu: s(illumination, k=5)", subtitle="Moon illumination model", tag="C")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.50, -0.25, 0, 0.25, 0.50), labels=c(0, 0.25, 0.50, 0.75, 1))+
  annotate("text", x=0.2, y=-0.18, label="Wt=0.09", size=5)
st_plot15

st_plot15 <- plot(msst15)[[1]]+
  geom_ribbon(data=disk_interval_st, fill="darkgray")+
  geom_ribbon(data=disk_interval_st2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="% Disk illumination", y="mu: s(illumination, k=5)", subtitle="Moon illumination model")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.50, -0.25, 0, 0.25, 0.50), labels=c(0, 0.25, 0.50, 0.75, 1))+
  annotate("text", x=0.2, y=-0.18, label="Wt=0.09", size=5)
st_plot15


# model 18
mest18 <- marginal_effects(model18st, effects="dewaters", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mest18)

msst18 <- marginal_smooths(model18st, resolution=1000)
msst18b <- marginal_smooths(model18st, resolution=1000, probs=c(0.1, 0.9))
msst18c <- marginal_smooths(model18st, resolution=1000, probs=c(0.25, 0.75))

dewater_interval_st <- msst18b$"mu: s(dewaters,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

dewater_interval_st2 <- msst18c$"mu: s(dewaters,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

st_plot18 <- plot(msst18)[[1]]+
  theme_bw()+
  geom_ribbon(data=dewater_interval_st, fill="darkgray")+
  geom_ribbon(data=dewater_interval_st2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="% Day dewatered", y="mu: s(dewater, k=5)", subtitle="Dewatered channel model", tag="E")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.004632069, 0.01536793, 0.03536793, 0.05536793, 0.07536793), labels=c(0, 0.02, 0.04, 0.06, 0.08))+
  annotate("text", x=0.055, y=-0.28, label="Wt=0.14", size=5)
st_plot18


# model 6
mest6 <- marginal_effects(model6st, effects="dlows", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mest6)

msst6 <- marginal_smooths(model6st, resolution=1000)
msst6b <- marginal_smooths(model6st, resolution=1000, probs=c(0.1, 0.9))
msst6c <- marginal_smooths(model6st, resolution=1000, probs=c(0.25, 0.75))

dlow_interval_st <- msst6b$"mu: s(dlows,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

dlow_interval_st2 <- msst6c$"mu: s(dlows,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

st_plot6 <- plot(msst6)[[1]]+
  theme_bw()+
  geom_ribbon(data=dlow_interval_st, fill="darkgray")+
  geom_ribbon(data=dlow_interval_st2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tidal low", y="mu: s(dlow, k=5)",  subtitle="Tidal low model", tag="D")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.2, 0, 0.2, 0.4), labels=c(0.8, 1.0, 1.2, 1.4))+
  annotate("text", x=0.2, y=-0.75, label="Wt=0.09", size=5)
st_plot6

st_plot6 <- plot(msst6)[[1]]+
  theme_bw()+
  geom_ribbon(data=dlow_interval_st, fill="darkgray")+
  geom_ribbon(data=dlow_interval_st2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tidal low", y="mu: s(dlow, k=5)",  subtitle="Tidal low model")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.2, 0, 0.2, 0.4), labels=c(0.8, 1.0, 1.2, 1.4))+
  annotate("text", x=0.2, y=-0.75, label="Wt=0.09", size=5)
st_plot6


# model 16
mest16 <- marginal_effects(model16st, effects="Twilight", robust=FALSE, re_formula=NA, resolution=1000, method = "fitted", probs = c(0.25,0.75))

st_plot16 <- plot(mest16)[[1]]+
  theme_bw()+
  labs(title="", subtitle="Day-night-twilight model", tag="B")+
  theme(text=element_text(family="Calibri", size=14), axis.title=element_blank(), 
        axis.text.x = element_text(angle=30, hjust=1, vjust=1))+
  annotate("text", x="Dawn", y=0.55, label="Wt=0.16", size=5)+
  ylim(0,0.6)
st_plot16

st_plot16 <- plot(mest16)[[1]]+
  theme_bw()+
  labs(title="", subtitle="Day-night-twilight model")+
  theme(text=element_text(family="Calibri", size=14), axis.title=element_blank(), 
        axis.text.x = element_text(angle=30, hjust=1, vjust=1))+
  annotate("text", x="Dawn", y=0.55, label="Wt=0.16", size=5)+
  ylim(0,0.6)
st_plot16


# model 21
mest21 <- marginal_effects(model21st, effects="phase8", robust=FALSE, re_formula=NA, resolution=1000, method = "fitted", probs = c(0.25,0.75))

st_plot21 <- plot(mest21)[[1]]+
  theme_bw()+
  labs(y="Detections", title="Splittail", subtitle="Lunar phase 8 model", tag="A")+
  theme(text=element_text(family="Calibri", size=14), axis.title.x=element_blank(), 
        axis.text.x = element_text(angle=30, hjust=1, vjust=1))+
  annotate("text", x="New", y=0.55, label="Wt=0.26", size=5)+
  ylim(0,0.6) #  
st_plot21


st_plot21 <- plot(mest21)[[1]]+
  theme_bw()+
  labs(y="Detections", title="Splittail", subtitle="Lunar phase 8 model")+
  theme(text=element_text(family="Calibri", size=14), axis.title.x=element_blank(), 
        axis.text.x = element_text(angle=90, hjust=1, vjust=1))+
  annotate("text", x="Waxing crescent", y=0.55, label="Wt=0.26", size=5)+
  ylim(0,0.6) #  
st_plot21

library(cowplot)
plot_grid(st_plot21, st_plot16, nrow=1, ncol=2, rel_widths = c(1,0.5))
plot_grid(st_plot15, st_plot6, st_plot18, nrow=1, ncol=3)


library(patchwork)
st_plot21 + st_plot16 + plot_layout(ncol=2, widths = c(1,0.65)) - {
  st_plot15 + st_plot6 + st_plot18 + plot_layout(ncol=3)
  } + 
  plot_layout(ncol=1)

# plot for manuscript
st_plot21 - {
  st_plot16 + st_plot15  + plot_layout(ncol=2)
} +
  plot_layout(ncol=1)
```

### ppc and coef tab
```{r}
model21st <- readRDS("Models/Splittail/brms_zip_21st_subset.rds")
summary(model21st)

ppst <- pp_check(model21st, type = "bars") +
  theme_bw()+
  labs(title="Splittail", subtitle="Phase8 model posterior check")+
  theme(text=element_text(size=16, family="Calibri"))
ppst

stanst <- stanplot(model21st)+
  theme_bw()+
  labs(title="Splittail", subtitle="Phase8 model coefficients")+
  theme(text=element_text(size=16, family="Calibri"))
stanst
```


## Striped bass

```{r}
model00sb <- readRDS("Models/Striper/brms_zip_00sb_subset.rds")
model0sb <- readRDS("Models/Striper/brms_zip_0sb_subset.rds")
model1sb <- readRDS("Models/Striper/brms_zip_1sb_subset.rds")
model2sb <- readRDS("Models/Striper/brms_zip_2sb_subset.rds")
model3sb <- readRDS("Models/Striper/brms_zip_3sb_subset.rds")
model4sb <- readRDS("Models/Striper/brms_zip_4sb_subset.rds")
model5sb <- readRDS("Models/Striper/brms_zip_5sb_subset.rds")
model6sb <- readRDS("Models/Striper/brms_zip_6sb_subset.rds")
model7sb <- readRDS("Models/Striper/brms_zip_7sb_subset.rds")
model8sb <- readRDS("Models/Striper/brms_zip_8sb_subset.rds")
model9sb <- readRDS("Models/Striper/brms_zip_9sb_subset.rds")
model10sb <- readRDS("Models/Striper/brms_zip_10sb_subset.rds")
model11sb <- readRDS("Models/Striper/brms_zip_11sb_subset.rds")
model12sb <- readRDS("Models/Striper/brms_zip_12sb_subset.rds")
model13sb <- readRDS("Models/Striper/brms_zip_13sb_subset.rds")
model14sb <- readRDS("Models/Striper/brms_zip_14sb_subset.rds")
model15sb <- readRDS("Models/Striper/brms_zip_15sb_subset.rds")
model16sb <- readRDS("Models/Striper/brms_zip_16sb_subset.rds")
model17sb <- readRDS("Models/Striper/brms_zip_17sb_subset.rds")
model18sb <- readRDS("Models/Striper/brms_zip_18sb_subset.rds")
```

```{r}
# check for problematic observations: pareto k estimates (k > 0.7)

LOO(model00sb) # OK
LOO(model0sb) # OK
LOO(model1sb) # OK
LOO(model2sb) # OK
LOO(model3sb) # OK
LOO(model4sb) # 1 bad
LOO(model5sb) # OK
LOO(model6sb) # OK
LOO(model7sb) # OK
LOO(model8sb) # OK
LOO(model9sb) # OK
LOO(model10sb) # OK
LOO(model11sb) # OK
LOO(model12sb) # OK
LOO(model13sb) # OK
LOO(model14sb) # OK
LOO(model15sb) # OK
LOO(model16sb) # OK
LOO(model17sb) # OK
LOO(model18sb) # OK
LOO(model19sb) # OK
LOO(model20sb) # OK
LOO(model21sb) # OK

# striped bass reloo models if needed:

# conduct a reloo analysis and store as loo in the model object
# code updated with feedback from paul buerkner (brms author)
# https://discourse.mc-stan.org/t/model-weights-function-with-reloo/8643

# model 4
# add loo to model
model4sb$loo <- LOO(model4sb, reloo = TRUE, pointwise = TRUE, k_threshold=0.7)

# save as reloo model name
model4sb %>% saveRDS("Models/Striper/brms_zip_model4sb_reloo.rds")

model4sb_reloo <- readRDS("Models/Striper/brms_zip_model4sb_reloo.rds")
```

Add loo criterion
```{r}
model00sb <- add_loo(model00sb)
model0sb <- add_loo(model0sb)
model1sb <- add_loo(model1sb)
model2sb <- add_loo(model2sb)
model3sb <- add_loo(model3sb)
model4sb <- add_loo(model4sb)
model5sb <- add_loo(model5sb)
model6sb <- add_loo(model6sb)
model7sb <- add_loo(model7sb)
model8sb <- add_loo(model8sb)
model9sb <- add_loo(model9sb)
model10sb <- add_loo(model10sb)
model11sb <- add_loo(model11sb)
model12sb <- add_loo(model12sb)
model13sb <- add_loo(model13sb)
model14sb <- add_loo(model14sb)
model15sb <- add_loo(model15sb)
model16sb <- add_loo(model16sb)
```

```{r}
loo(model00sb)
loo(model0sb)
loo(model1sb)
loo(model2sb)
loo(model3sb)
loo(model4sb_reloo)
loo(model5sb)
loo(model6sb)
loo(model7sb)
loo(model8sb)
loo(model9sb)
loo(model10sb)
loo(model11sb)
loo(model12sb)
loo(model13sb)
loo(model14sb)
loo(model15sb)
loo(model16sb)
loo(model17sb)
loo(model18sb)
loo(model19sb)
```

### model results
```{r}
# brms compare loo
sb_loo <- loo_compare(model00sb, model0sb, model1sb, model2sb, model3sb, model4sb_reloo, model5sb, model6sb, model7sb, model8sb, model9sb, model10sb, model11sb, model12sb, model13sb, model14sb, model15sb, model16sb, model17sb, model18sb, model19sb, model20sb, model21sb)

sb_loo_df <- print(sb_loo, simplify=F, digits=1)
sb_loo_df

sb_loo_df2 <- sb_loo_df %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(model=rowname)
View(sb_loo_df2)

sb_loo2 <- cbind(loo_diff = sb_loo[, 1] * -2,
      se        = sb_loo[, 2] * 2) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(model=rowname)
View(sb_loo2)

sb_loo2 %>% write_csv("Data_output/STB_looic.csv")

# model weights
# regular weights
sb_mw_loo <- model_weights(model00sb, model0sb, model1sb, model2sb, model3sb, model4sb_reloo, model5sb, model6sb, model7sb, model8sb, model9sb, model10sb, model11sb, model12sb, model13sb, model14sb, model15sb, model16sb, model17sb, model18sb, model19sb, model20sb, model21sb, weights = "loo")
View(sb_mw_loo) 

sb_mw_loo <- data.frame(sb_mw_loo) %>% rownames_to_column() %>% 
  rename(model=rowname, model_weights=sb_mw_loo) %>% 
  mutate(model_weights=round(model_weights, digits=2)) %>% 
  arrange(desc(model_weights))
sb_mw_loo
View(sb_mw_loo)

sb_mw_loo %>% write_csv("Data_output/STB_Model_weights_loo.csv")

# stacked weights
sb_mw_loo2 <- model_weights(model00sb, model0sb, model1sb, model2sb, model3sb, model4sb_reloo, model5sb, model6sb, model7sb, model8sb, model9sb, model10sb, model11sb, model12sb, model13sb, model14sb, model15sb, model16sb, model17sb, model18sb, model19sb, model20sb, model21sb, weights = "loo2") # loo2 for Bayesian stacked weights
View(sb_mw_loo2) 

sb_mw_loo2 <- data.frame(sb_mw_loo2) %>% rownames_to_column() %>% 
  rename(model=rowname, stacked_weights=sb_mw_loo2) %>% 
  mutate(stacked_weights=round(stacked_weights, digits=2)) %>% 
  arrange(desc(stacked_weights))
sb_mw_loo2
View(sb_mw_loo2)

sb_mw_loo2 %>% write_csv("Data_output/STB_Stacked_weights_loo2.csv")

# create a table for manuscript
sb_tbl_loo <- sb_loo_df2 %>% #looic
  left_join(sb_loo2, by="model") %>% 
  left_join(sb_mw_loo, by="model") %>% # regular model weights
  left_join(sb_mw_loo2, by="model") # stacked model weights
View(sb_tbl_loo)

sb_tbl_loo %>% write_csv("Data_output/STB_Model_looic_compare_loo_vs_loo2.csv")
```

### effect plots
```{r}
# marginal effects and splines plots
# model 1
mesb1 <- marginal_effects(model1sb, effects="stage_z", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mesb1)

mssb1 <- marginal_smooths(model1sb, resolution=1000, probs=c(0.025, 0.975))
mssb1b <- marginal_smooths(model1sb, resolution=1000, probs=c(0.1, 0.9))
mssb1c <- marginal_smooths(model1sb, resolution=1000, probs=c(0.25, 0.75))

depth_interval_sb <- mssb1b$"mu: s(stage_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

depth_interval_sb2 <- mssb1c$"mu: s(stage_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

sb_plot1 <- plot(mssb1)[[1]]+
  geom_ribbon(data=depth_interval_sb, fill="darkgray")+
  geom_ribbon(data=depth_interval_sb2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tide height (z-score transformed)", y="mu: s(theight, k=5)", title="Striped Bass", subtitle="Tide height model", tag="A")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=1.2, y=-0.27, label="Wt=0.26", size=5)
sb_plot1


sb_plot1 <- plot(mssb1)[[1]]+
  geom_ribbon(data=depth_interval_sb, fill="darkgray")+
  geom_ribbon(data=depth_interval_sb2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Water depth (z-score transformed)", y="mu: s(depth, k=5)", title="Striped Bass", subtitle="Water depth model")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=1.2, y=-0.27, label="Wt=0.26", size=5)
sb_plot1


# model 6
mesb6 <- marginal_effects(model6sb, effects="dlows", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mesb6)

mssb6 <- marginal_smooths(model6sb, resolution=1000)
mssb6b <- marginal_smooths(model6sb, resolution=1000, probs=c(0.1, 0.9))
mssb6c <- marginal_smooths(model6sb, resolution=1000, probs=c(0.25, 0.75))

dlow_interval_sb <- mssb6b$"mu: s(dlows,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

dlow_interval_sb2 <- mssb6c$"mu: s(dlows,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

sb_plot6 <- plot(mssb6)[[1]]+
  theme_bw()+
  geom_ribbon(data=dlow_interval_sb, fill="darkgray")+
  geom_ribbon(data=dlow_interval_sb2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tidal low", y="mu: s(tlow, k=5)", subtitle="Tidal low model", tag="C")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.2, 0, 0.2, 0.4), labels=c(0.8, 1.0, 1.2, 1.4))+
  annotate("text", x=0.35, y=-1.1, label="Wt=0.18", size=5)
sb_plot6


sb_plot6 <- plot(mssb6)[[1]]+
  theme_bw()+
  geom_ribbon(data=dlow_interval_sb, fill="darkgray")+
  geom_ribbon(data=dlow_interval_sb2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tidal low", y="mu: s(dlow, k=5)", subtitle="Tidal low model")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.2, 0, 0.2, 0.4), labels=c(0.8, 1.0, 1.2, 1.4))+
  annotate("text", x=0.35, y=-1.1, label="Wt=0.18", size=5)
sb_plot6


# model 8
mesb8 <- marginal_effects(model8sb, effects="dinequalitys", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mesb8)

mssb8 <- marginal_smooths(model8sb, resolution=1000)
mssb8b <- marginal_smooths(model8sb, resolution=1000, probs=c(0.1, 0.9))
mssb8c <- marginal_smooths(model8sb, resolution=1000, probs=c(0.25, 0.75))

din_interval_sb <- mssb8b$"mu: s(dinequalitys,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

din_interval_sb2 <- mssb8c$"mu: s(dinequalitys,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

sb_plot8 <- plot(mssb8)[[1]]+
  theme_bw()+
  geom_ribbon(data=din_interval_sb, fill="darkgray")+
  geom_ribbon(data=din_interval_sb2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tidal inequality", y="mu: s(tinequality, k=5)", subtitle="Tidal inequality model", tag="D")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.6, -0.4, -0.2, 0, 0.2, 0.4), labels=c(-0.25, 0, 0.25, 0.50, 0.75, 1))+
  annotate("text", x=0.2, y=-0.45, label="Wt=0.13", size=5)
sb_plot8

sb_plot8 <- plot(mssb8)[[1]]+
  theme_bw()+
  geom_ribbon(data=din_interval_sb, fill="darkgray")+
  geom_ribbon(data=din_interval_sb2, fill="gray50")+
  geom_line(size=1, col="black")+
    geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tidal inequality", y="mu: s(dinequality, k=5)", subtitle="Tidal inequality model")+
  theme(text=element_text(family="Calibri", size=14))+
  scale_x_continuous(breaks=c(-0.6, -0.4, -0.2, 0, 0.2, 0.4), labels=c(-0.25, 0, 0.25, 0.50, 0.75, 1))+
  annotate("text", x=0.2, y=-0.45, label="Wt=0.13", size=5)
sb_plot8


# model 14
mesb14 <- marginal_effects(model14sb, effects="temp_z", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(mesb14)

mssb14 <- marginal_smooths(model14sb, resolution=1000)
mssb14b <- marginal_smooths(model14sb, resolution=1000, probs=c(0.1, 0.9))
mssb14c <- marginal_smooths(model14sb, resolution=1000, probs=c(0.25, 0.75))

temp_interval_sb <- mssb14b$"mu: s(temp_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

temp_interval_sb2 <- mssb14c$"mu: s(temp_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

xlab <- "Temperature (z-score transformed)"

sb_plot14 <- plot(mssb14)[[1]]+
  theme_bw()+
  geom_ribbon(data=temp_interval_sb, fill="darkgray")+
  geom_ribbon(data=temp_interval_sb2, fill="gray50")+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x=xlab, y="mu: s(temp, k=5)", subtitle="Temperature model", tag="B")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=2.5, y=-0.45, label="Wt=0.09", size=5)
sb_plot14



# model 16
mesb16 <- marginal_effects(model16sb, effects="Twilight", robust=FALSE, re_formula=NA, resolution=1000, method = "fitted", probs=c(0.25,0.75))

sb_plot16 <- plot(mesb16)[[1]]+
  theme_bw()+
  labs(y="Detections", subtitle="Day-night-twilight model", tag="E")+
  theme(text=element_text(family="Calibri", size=14), axis.title.x = element_blank(), 
        axis.text.x = element_text(angle=30, hjust=1, vjust=1))+
  annotate("text", x="Dawn", y=0.28, label="Wt=0.10", size=5)+
  ylim(0,0.3)
sb_plot16

library(patchwork)
sb_plot1 + sb_plot14 + plot_layout(ncol=2) -  {
  sb_plot6 + sb_plot8 + plot_layout(ncol=2)
} + 
  plot_layout(nrow=2) - {
    sb_plot16
  } + 
  plot_layout(nrow=2, heights=c(1,0.4))
```

### ppc and coef tab
```{r}
model1sb <- readRDS("Models/Striper/brms_zip_1sb_subset.rds")
summary(model1sb)

ppsb <- pp_check(model1sb, type = "bars") +
  theme_bw()+
  labs(title="Striped Bass", subtitle="Tide height model posterior check")+
  theme(text=element_text(size=16, family="Calibri"))
ppsb

stansb <- stanplot(model1sb)+
  theme_bw()+
  labs(title="Striped Bass", subtitle="Tide height model coefficients")+
  theme(text=element_text(size=16, family="Calibri"))
stansb
```


## Tule perch
```{r}
# tule perch
model00tp <- readRDS("Models/Tuleperch/brms_zip_00tp_subset.rds")
model0tp <- readRDS("Models/Tuleperch/brms_zip_0tp_subset.rds")
model1tp <- readRDS("Models/Tuleperch/brms_zip_1tp_subset.rds")
model2tp <- readRDS("Models/Tuleperch/brms_zip_2tp_subset.rds")
model3tp <- readRDS("Models/Tuleperch/brms_zip_3tp_subset.rds")
model4tp <- readRDS("Models/Tuleperch/brms_zip_4tp_subset.rds")
model5tp <- readRDS("Models/Tuleperch/brms_zip_5tp_subset.rds")
model6tp <- readRDS("Models/Tuleperch/brms_zip_6tp_subset.rds")
model7tp <- readRDS("Models/Tuleperch/brms_zip_7tp_subset.rds")
model8tp <- readRDS("Models/Tuleperch/brms_zip_8tp_subset.rds")
model9tp <- readRDS("Models/Tuleperch/brms_zip_9tp_subset.rds")
model10tp <- readRDS("Models/Tuleperch/brms_zip_10tp_subset.rds")
model11tp <- readRDS("Models/Tuleperch/brms_zip_11tp_subset.rds")
model12tp <- readRDS("Models/Tuleperch/brms_zip_12tp_subset.rds")
model13tp <- readRDS("Models/Tuleperch/brms_zip_13tp_subset.rds")
model14tp <- readRDS("Models/Tuleperch/brms_zip_14tp_subset.rds")
model15tp <- readRDS("Models/Tuleperch/brms_zip_15tp_subset.rds")
model16tp <- readRDS("Models/Tuleperch/brms_zip_16tp_subset.rds")
model17tp <- readRDS("Models/Tuleperch/brms_zip_17tp_subset.rds")
model18tp <- readRDS("Models/Tuleperch/brms_zip_18tp_subset.rds")
```

```{r}
# check for problematic observations: pareto k estimates (k > 0.7)

LOO(model00tp) # OK
LOO(model0tp) # OK
LOO(model1tp) # OK
LOO(model2tp) # OK
LOO(model3tp) # OK
LOO(model4tp) # OK
LOO(model5tp) # OK
LOO(model6tp) # OK
LOO(model7tp) # OK
LOO(model8tp) # OK
LOO(model9tp) # OK
LOO(model10tp) # OK
LOO(model11tp) # OK
LOO(model12tp) # OK
LOO(model13tp) # OK
LOO(model14tp) # OK
LOO(model15tp) # OK
LOO(model16tp) # OK
LOO(model17tp) # OK
LOO(model18tp) # OK
LOO(model19tp) # OK
LOO(model20tp) # OK
LOO(model21tp) # OK
```

Add loo criterion
```{r}
model0000tp <- add_loo(model0000tp)
model000tp <- add_loo(model000tp)
model00tp <- add_loo(model00tp)
model0tp <- add_loo(model0tp)
model1tp <- add_loo(model1tp)
model2tp <- add_loo(model2tp)
model3tp <- add_loo(model3tp)
model4tp <- add_loo(model4tp)
model5tp <- add_loo(model5tp)
model6tp <- add_loo(model6tp)
model7tp <- add_loo(model7tp)
model8tp <- add_loo(model8tp)
model9tp <- add_loo(model9tp)
model10tp <- add_loo(model10tp)
model11tp <- add_loo(model11tp)
model12tp <- add_loo(model12tp)
model13tp <- add_loo(model13tp)
model14tp <- add_loo(model14tp)
model15tp <- add_loo(model15tp)
model16tp <- add_loo(model16tp)
model17tp <- add_loo(model17tp)
model18tp <- add_loo(model18tp)
model19tp <- add_loo(model19tp)
model20tp <- add_loo(model20tp)

summary(model000tp)
```

### model results
```{r}
# brms compare loo
tp_loo <- loo_compare(model00tp, model0tp, model1tp, model2tp, model3tp, model4tp, model5tp, model6tp, model7tp, model8tp, model9tp, model10tp, model11tp, model12tp, model13tp, model14tp, model15tp, model16tp, model17tp, model18tp, model19tp, model20tp, model21tp)

tp_loo_df <- print(tp_loo, simplify=F, digits=1)
tp_loo_df

tp_loo_df2 <- tp_loo_df %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(model=rowname)
View(tp_loo_df2)

tp_loo2 <- cbind(loo_diff = tp_loo[, 1] * -2,
      se        = tp_loo[, 2] * 2) %>% 
  as.data.frame() %>% 
  rownames_to_column() %>% 
  rename(model=rowname)
View(tp_loo2)

# fix this file path - sub redo v10
tp_loo2 %>% write_csv("Data_output/TUP_looic.csv")

# model weights
# regular weights
tp_mw_loo <- model_weights(model00tp, model0tp, model1tp, model2tp, model3tp, model4tp, model5tp, model6tp, model7tp, model8tp, model9tp, model10tp, model11tp, model12tp, model13tp, model14tp, model15tp, model16tp, model17tp, model18tp, model19tp, model20tp, model21tp, weights = "loo") 
View(tp_mw_loo) 

tp_mw_loo <- data.frame(tp_mw_loo) %>% rownames_to_column() %>% 
  rename(model=rowname, model_weights=tp_mw_loo) %>% 
  mutate(model_weights=round(model_weights, digits=2)) %>% 
  arrange(desc(model_weights))
tp_mw_loo
View(tp_mw_loo)

tp_mw_loo %>% write_csv("Data_output/TUP_Model_weights_loo.csv")

# stacked weights
tp_mw_loo2 <- model_weights(model00tp, model0tp, model1tp, model2tp, model3tp, model4tp, model5tp, model6tp, model7tp, model8tp, model9tp, model10tp, model11tp, model12tp, model13tp, model14tp, model15tp, model16tp, model17tp, model18tp, model19tp, model20tp, model21tp, weights = "loo2") # loo2 for stacked weights
View(tp_mw_loo2) 

tp_mw_loo2 <- data.frame(tp_mw_loo2) %>% rownames_to_column() %>% 
  rename(model=rowname, stacked_weights=tp_mw_loo2) %>% 
  mutate(stacked_weights=round(stacked_weights, digits=2)) %>% 
  arrange(desc(stacked_weights))
tp_mw_loo2
View(tp_mw_loo2)

tp_mw_loo2 %>% write_csv("Data_output/TUP_Stacked_weights_loo2.csv")

# create a table for manuscript
tp_tbl_loo <- tp_loo_df2 %>% #looic
  left_join(tp_mw_loo, by="model") %>% # regular model weights
  left_join(tp_mw_loo2, by="model") # stacked model weights
View(tp_tbl_loo)

tp_tbl_loo %>% write_csv("Data_output/TUP_Model_looic_compare_loo_vs_loo2.csv")
```

### effect plots
```{r}

# marginal effects and splines plots
# model 9
metp9 <- marginal_effects(model9tp, effects="slopes", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(metp9)

mstp9 <- marginal_smooths(model9tp, resolution=1000)
mstp9b <- marginal_smooths(model9tp, resolution=1000, probs=c(0.1,0.9))
mstp9c <- marginal_smooths(model9tp, resolution=1000, probs=c(0.25,0.75))

slope_interval_tp <- mstp9b$"mu: s(slopes,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

slope_interval_tp2 <- mstp9c$"mu: s(slopes,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

tp_plot9 <- plot(mstp9)[[1]]+
  geom_ribbon(data=slope_interval_tp, fill="darkgray")+
  geom_ribbon(data=slope_interval_tp2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Rate of change in tide height", y="mu: s(trate, k=5)", title="Tule Perch", subtitle="Rate of change in tide height model", tag="A")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=9, y=-0.1, label="Wt=0.26", size=5)
tp_plot9


tp_plot9 <- plot(mstp9)[[1]]+
  geom_ribbon(data=slope_interval_tp, fill="darkgray")+
  geom_ribbon(data=slope_interval_tp2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Rate of change in water depth", y="mu: s(rate, k=5)", title="Tule Perch", subtitle="Rate of change in water depth model")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=9, y=-0.1, label="Wt=0.26", size=5)
tp_plot9


# model 1
metp1 <- marginal_effects(model1tp, effects="stage_z", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(metp1)

mstp1 <- marginal_smooths(model1tp, resolution=1000, probs=c(0.025, 0.975))
mstp1b <- marginal_smooths(model1tp, resolution=1000, probs=c(0.1, 0.9))
mstp1c <- marginal_smooths(model1tp, resolution=1000, probs=c(0.25, 0.75))

depth_interval_tp <- mstp1b$"mu: s(stage_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

depth_interval_tp2 <- mstp1c$"mu: s(stage_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

tp_plot1 <- plot(mstp1)[[1]]+
  geom_ribbon(data=depth_interval_tp, fill="darkgray")+
  geom_ribbon(data=depth_interval_tp2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Tide height (z-score transformed)", y="mu: s(theight, k=5)", subtitle="Tide height model", tag="B")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=1.2, y=-0.7, label="Wt=0.22", size=5)
tp_plot1


tp_plot1 <- plot(mstp1)[[1]]+
  geom_ribbon(data=depth_interval_tp, fill="darkgray")+
  geom_ribbon(data=depth_interval_tp2, fill="gray50")+
  theme_bw()+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x="Water depth (z-score transformed)", y="mu: s(depth, k=5)", subtitle="Water depth model")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=1.2, y=-0.7, label="Wt=0.22", size=5)
tp_plot1


# model 14
metp14 <- marginal_effects(model14tp, effects="temp_z", robust=TRUE, re_formula=NULL, resolution=1000, method = "fitted")
plot(metp14)

mstp14 <- marginal_smooths(model14tp, resolution=1000)
mstp14b <- marginal_smooths(model14tp, resolution=1000, probs=c(0.1, 0.9))
mstp14c <- marginal_smooths(model14tp, resolution=1000, probs=c(0.25, 0.75))

temp_interval_tp <- mstp14b$"mu: s(temp_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

temp_interval_tp2 <- mstp14c$"mu: s(temp_z,k=5)" %>% 
  select(`lower__`, `upper__`, `effect1__`, `estimate__`)

xlab <- "Temperature (z-score transformed)"

tp_plot14 <- plot(mstp14)[[1]]+
  theme_bw()+
  geom_ribbon(data=temp_interval_tp, fill="darkgray")+
  geom_ribbon(data=temp_interval_tp2, fill="gray50")+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x=xlab, y="mu: s(temp, k=5)", subtitle="Temperature model", tag="C")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=2.5, y=-0.6, label="Wt=0.16", size=5)
tp_plot14


tp_plot14 <- plot(mstp14)[[1]]+
  theme_bw()+
  geom_ribbon(data=temp_interval_tp, fill="darkgray")+
  geom_ribbon(data=temp_interval_tp2, fill="gray50")+
  geom_line(size=1, col="black")+
  geom_hline(yintercept = 0, lty=2, col="red")+
  labs(x=xlab, y="mu: s(temp, k=5)", subtitle="Temperature model")+
  theme(text=element_text(family="Calibri", size=14))+
  annotate("text", x=2.5, y=-0.6, label="Wt=0.16", size=5)
tp_plot14

tp_plot9 + tp_plot1 + tp_plot14 + plot_layout(nrow=1, ncol=3)
```


### ppc and coef tab
```{r}
model0tp <- readRDS("Models/Tuleperch/brms_zip_0tp_subset.rds")
summary(model0tp)

# model 0
pptp <- pp_check(model0tp, type = "bars") +
  theme_bw()+
  labs(title="Tule Perch", subtitle="Month model posterior check")+
  theme(text=element_text(size=16, family="Calibri"))
pptp

stantp <- stanplot(model0tp)+
  theme_bw()+
  labs(title="Tule Perch", subtitle="Month model coefficients")+
  theme(text=element_text(size=16, family="Calibri"))
stantp
```

# model checking all spp

## ppc and stan plots
```{r}
library(patchwork)
ppsb + ppst + pptp + plot_layout(nrow=3, ncol=1)

stansb + stanst + stantp + plot_layout(nrow=3, ncol=1)
```



Extra stuff I tried for model averaging but didn't end up using.
```{r}
(tp_nd <- unique(dtp[, c("PIT", "logdays", "logdet", "logmin", "mnth", "PIT", "stage_z", "tide", "slopes", "temp_z")]))

# compute model-averaged predicted values
tp_pp_avg <- pp_average(model0tp, model1tp, model4tp, model9tp, model14tp, weights = "loo2", method = "predict", summary = TRUE, 
  probs = c(0.025, 0.975), newdata = tp_nd)
View(tp_pp_avg)

# compute model-averaged fitted values
tp_pp_avg_fit <- pp_average(model0tp, model1tp, model4tp, model9tp, model14tp, weights = "loo2", method = "fitted", summary = TRUE, 
  probs = c(0.025, 0.975), newdata = tp_nd)
View(tp_pp_avg_fit)

tp_fit <- as_tibble(tp_pp_avg_fit)
tp_predict <- as_tibble(tp_pp_avg)

ggplot(tp_predict, aes(Estimate))+
  geom_density(fill="darkblue", alpha=0.5)+
  geom_density(data=tp_fit, aes(Estimate), fill="lightblue", alpha=0.7)+
  labs(title= "Tule Perch", subtitle="Model-averaged posterior predictive distribution")+
  xlim(0,8)
```
