---
title: "Ch1_Abun_Biom_Growth"
author: "Denise Colombano"
date: "March 10, 2017"
output: html_document
---
## Purpose of this exercise
Here I will compare abundance, biomass and growth of YOY fish to answer the following questions:

(1) Where are hotspots of abundance and biomass?
      - Identify and classify nurseries using plots of mean annual abundance and biomass per slough and/or per year
(2) When is peak abundance occurring? When is peak growth occurring?
      - Calculate change in mean biomass (=growth) on monthly and annual basis
      
Note: Better to answer Question 2 first in order to figure out which months (ex 5-10?) to include in future analyses.

```{r, load libraries and data}
# load libraries
library(tidyverse)
library(ggplot2)
library(magrittr)
library(ggthemes)

chapter1 <- read_csv("~/GitHub/suisunmarsh/Ch1/Data/Data_Ch1_Abundance_1995-2016_Redo.txt")
chapter1_biom2 <- read_csv("~/GitHub/suisunmarsh/Ch1/Data_output/Data_Ch1_BiomassEstimates_1995-2016_Redo.csv")

# check DFs for NAs or other import errors
summary(chapter1)
summary(chapter1_biom2)


```


Approach 1. Calculate sums and averages within equations using raw abundance/biomass, rather than CPUE and GPUE.

### QUESTION 1
## HOTSPOT: ABUNDANCE
# Borrow code from previous analysis in file "Ch1_Abundance" to look at pos/neg differences in abundance
```{r}
## tidy dataset
abundance <- chapter1_biom2 %>% 
  select(Slough, YearMo, Mo, Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  filter(Mo >4) %>% 
  filter(Mo < 10) %>% 
  group_by(Slough, Yr) %>% 
  summarize(st0_cpue = sum(`ST_Age-0`/ TowMin), sb0_cpue = sum(`SB_Age-0`/ TowMin), tp0_cpue = sum(`TP_Age-0`/ TowMin))

abundance[is.na(abundance)] <- 0

summary(abundance)
View(abundance)

# Calculate annual marsh-wide average (no slough)
mean_abundance <- abundance %>%
  group_by(Yr) %>% 
  summarize(st0_mean = mean(st0_cpue), sb0_mean = mean(sb0_cpue), tp0_mean = mean(tp0_cpue))
View(mean_abundance)
summary(mean_abundance)
```

```{r, join back with chapter1 dataframe}

# Following this lesson:

## **https://ismayc.github.io/moderndiver-book/5-manip.html#joining-data-frames**

# inner join
abundance_joined <- abundance %>%
  inner_join(mean_abundance, by="Yr")
summary(abundance_joined)
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
abundance_differences <- abundance_joined %>% 
  mutate(SB = sb0_cpue - sb0_mean, ST = st0_cpue - st0_mean, TP = tp0_cpue - tp0_mean) %>%
  select(Slough, Yr, SB, ST, TP) %>% 
  glimpse

a <- abundance_differences # shorthand
View(a)
summary(a)
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
a$Slough <- as.factor(a$Slough)
a$Yr <- as.factor(a$Yr)

# tidyr gather
a_gathered <- a %>% 
  gather(Species, Relative_CPUE, SB:TP)
View(a_gathered)
```

Plot relative differences:
First let's look at disproportionately greater than average cPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, mean difference bar charts}
# with legend
a_gathered$Above_Annual_Mean <- a_gathered$Relative_CPUE >= 0
bar <- ggplot(a_gathered, aes(x=Yr, y=Relative_CPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity") +
  theme_economist_white() +
  scale_fill_economist()+
  guides(fill=FALSE) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Catch Per Minute (Annual Mean=0)", x="") #title="Relative Annual Catch of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=14), axis.title.y = element_text(size=12), title = element_text(size=10), axis.line.y = element_line(size=0.5), strip.text = element_text(size=10))

# alternative bar chart
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=16), axis.title.y = element_text(size=16))
```


#Approach 2. Use CPUE and GPUE values in the "Chapter1_biom2" DF instead of calculating from raw abundance.

```{r}
# Calculate mean at each slough and year combination
abundance2 <- chapter1_biom2 %>% 
  select(Slough, YearMo, YrSlough, Mo, Yr, ST0_CPUE, SB0_CPUE, TP0_CPUE) %>% 
  filter(Mo>4 | Mo<10) %>% 
  group_by(Slough, Yr) %>% 
  summarize(st0_cpue = sum(ST0_CPUE), sb0_cpue = sum(SB0_CPUE), tp0_cpue = sum(TP0_CPUE))
summary(abundance2)
View(abundance2)

# Calculate annual marsh-wide average
mean_abundance2 <- abundance2 %>%
  select(Yr, st0_cpue, sb0_cpue, tp0_cpue) %>% 
  group_by(Yr) %>% 
  summarize(st0_mean = mean(st0_cpue), sb0_mean = mean(sb0_cpue), tp0_mean = mean(tp0_cpue))
View(mean_abundance2)
summary(mean_abundance2)

abundance_joined2 <- abundance2 %>%
  inner_join(mean_abundance2, by="Yr")
summary(abundance_joined2)
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
abundance_differences2 <- abundance_joined2 %>% 
  mutate(Splittail = st0_cpue - st0_mean, StripedBass = sb0_cpue - sb0_mean, TulePerch = tp0_cpue - tp0_mean) %>%
  select(Slough, Yr, Splittail, StripedBass, TulePerch) %>% 
  glimpse

a2 <- abundance_differences2 # shorthand
View(a2)
summary(a2)
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
a2$Slough <- as.factor(a2$Slough)
a2$Yr <- as.factor(a2$Yr)

# tidyr gather
a_gathered2 <- a2 %>% 
  gather(Species, Relative_CPUE, Splittail:TulePerch)
View(a_gathered2)
```

Plot relative differences:
First let's look at disproportionately greater than average cPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, mean difference bar charts}
# with legend
a_gathered2$Above_Annual_Mean <- a_gathered2$Relative_CPUE >= 0
bar <- ggplot(a_gathered2, aes(x=Yr, y=Relative_CPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity") +
  theme_economist_white() +
  scale_fill_economist()+
  guides(fill=FALSE) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Catch Per Minute (Annual Mean=0)", x="") #, title="Relative Annual Catch of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=16), axis.title.y = element_text(size=16), title = element_text(size=20), axis.line.y = element_line(size=0.5), strip.text = element_text(size=14))

# alternative bar chart
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=16), axis.title.y = element_text(size=16))


```




## HOTSPOT: BIOMASS

```{r}
## tidy dataset
biomass <- chapter1_biom2 %>% 
  select(Slough, YearMo, Mo, Yr, st0_b_tot, sb0_b_tot, tp0_b_tot, TowMin) %>% 
  filter(Mo>4 ) %>% 
  filter(Mo<10) %>% 
  group_by(Slough, Yr) %>% 
  summarize(st0_gpue = sum(st0_b_tot/ TowMin), sb0_gpue = sum(sb0_b_tot/ TowMin), tp0_gpue = sum(tp0_b_tot/ TowMin))

biomass[is.na(biomass)] <- 0

summary(biomass)

# Group by and Summarize
mean_biomass <- biomass %>%
  group_by(Yr) %>% 
  summarize(st0_mean = mean(st0_gpue), sb0_mean = mean(sb0_gpue), tp0_mean = mean(tp0_gpue))
View(mean_biomass)
summary(mean_biomass)
```

```{r, join back with chapter1 dataframe}

# Following this lesson:

## **https://ismayc.github.io/moderndiver-book/5-manip.html#joining-data-frames**

# inner join
biomass_joined <- biomass %>%
  inner_join(mean_biomass, by="Yr")
summary(biomass_joined)
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
biomass_differences <- biomass_joined %>% 
  mutate(Splittail = st0_gpue - st0_mean, Striped_Bass = sb0_gpue - sb0_mean, Tule_Perch = tp0_gpue - tp0_mean) %>%
  select(Slough, Yr, Splittail, Striped_Bass, Tule_Perch) %>% 
  glimpse

b <- biomass_differences # shorthand
View(b)
summary(b)
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
b$Slough <- as.factor(b$Slough)
b$Yr <- as.factor(b$Yr)

# tidyr gather
b_gathered <- b %>% 
  gather(Species, Relative_GPUE, Splittail:Tule_Perch)
View(b_gathered)
```

Plot relative differences:
First let's look at disproportionately greater than average GPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, mean difference bar charts}
# with legend
b_gathered$Above_Annual_Mean <- b_gathered$Relative_GPUE >= 0

bar <- ggplot(b_gathered, aes(x=Yr, y=Relative_GPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity") +
  theme_economist_white() +
  scale_fill_economist()+
  guides(fill=FALSE) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Grams Per Minute (Annual Mean=0)", x="") #title="Relative Annual Biomass of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=14), axis.title.y = element_text(size=12), title = element_text(size=10), axis.line.y = element_line(size=0.5), strip.text = element_text(size=10))

# alternative bar chart
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=16), axis.title.y = element_text(size=16))
```


#Approach 2. Use CPUE and GPUE values in the "Chapter1_biom2" DF instead of calculating from raw abundance.
```{r}
## tidy dataset
biomass2 <- chapter1_biom2 %>% 
  select(Slough, YearMo, Mo, Yr, st0_gpue, sb0_gpue, tp0_gpue) %>% 
  filter(Mo>4) %>% 
  filter(Mo<10) %>% 
  group_by(Slough, Yr) %>% 
  summarize(st0_gpue = sum(st0_gpue), sb0_gpue = sum(sb0_gpue), tp0_gpue = sum(tp0_gpue))
View(biomass2)

# Group by and Summarize
mean_biomass2 <- biomass2 %>%
  group_by(Yr) %>% 
  summarize(st0_mean = mean(st0_gpue), sb0_mean = mean(sb0_gpue), tp0_mean = mean(tp0_gpue))
View(mean_biomass2)
summary(mean_biomass2)
```

```{r, join back with chapter1 dataframe}

# Following this lesson:

## **https://ismayc.github.io/moderndiver-book/5-manip.html#joining-data-frames**

# inner join
biomass_joined2 <- biomass2 %>%
  inner_join(mean_biomass2, by="Yr")
summary(biomass_joined2)
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
biomass_differences2 <- biomass_joined2 %>% 
  mutate(Splittail = st0_gpue - st0_mean, Striped_Bass = sb0_gpue - sb0_mean, Tule_Perch = tp0_gpue - tp0_mean) %>%
  select(Slough, Yr, Splittail, Striped_Bass, Tule_Perch) %>% 
  glimpse

b2 <- biomass_differences2 # shorthand
View(b2)
summary(b2)
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
b2$Slough <- as.factor(b2$Slough)
b2$Yr <- as.factor(b2$Yr)

# tidyr gather
b_gathered2 <- b2 %>% 
  gather(Species, Relative_GPUE, Splittail:Tule_Perch)
View(b_gathered2)
```

Plot relative differences:
First let's look at disproportionately greater than average GPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, mean difference bar charts}
# with legend
b_gathered2$Above_Annual_Mean <- b_gathered2$Relative_GPUE >= 0

bar <- ggplot(b_gathered2, aes(x=Yr, y=Relative_GPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity") +
  theme_economist_white() +
  scale_fill_economist()+
  guides(fill=FALSE) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Grams Per Minute (Annual Mean=0)", x="", title="Relative Annual Biomass of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=14), axis.title.y = element_text(size=12), title = element_text(size=10), axis.line.y = element_line(size=0.5), strip.text = element_text(size=10))
```




## BAR CHART SUMMARIES: HOTSPOTS for ABUNDANCE & BIOMASS at EACH SLOUGH (AVERAGED OVER TOTAL ANNUAL)
# Borrow code from Ch1_Abundance file again

```{r, calculate average annual abundance over 21 years}
## Calculate average ANNUAL cPUE (21 data points)

abunbiom <- abundance %>%
  inner_join(biomass, by=c("Slough", "Yr")) %>% 
  select(Slough, Yr, YrSlough.x, st0_cpue:tp0_cpue, st0_gpue:tp0_gpue) %>% 
  rename(YrSlough=YrSlough.x)

abunbiom_gathered <- abunbiom %>%
  select(Slough, Yr, st0_cpue:tp0_cpue) %>% 
  rename(Splittail=st0_cpue, StripedBass=sb0_cpue, TulePerch=tp0_cpue) %>% 
  gather(Species, CPUE, Splittail:TulePerch) %>% 
  unite_("YrSlough", c("Yr","Slough"))

?unite

abunbiom_gathered2 <- abunbiom %>% 
  select(Slough, Yr, st0_gpue:tp0_gpue) %>% 
  rename(Splittail=st0_gpue, StripedBass=sb0_gpue, TulePerch=tp0_gpue) %>% 
  gather(Species, GPUE, Splittail:TulePerch) %>% 
  unite_("YrSlough", c("Yr","Slough"))

abunbiom_joined <- abunbiom_gathered %>% 
  inner_join(abunbiom_gathered2, by=c("YrSlough", "Species")) 

abunbiom_joined2 <- abunbiom_joined %>% 
  write_csv("~/GitHub/suisunmarsh/Ch1/Data_Ch1_Abun_Biom_YrSlough.csv")


# calculate mean CPUE and SE per slough
# use plotrix pacakge to calculate SE
annual_mean_cpue <- abundance %>% 
  group_by(Slough) %>% 
  summarize(Splittail = mean(st0_cpue), StripedBass = mean(sb0_cpue), TulePerch = mean(tp0_cpue), Splittail_SE =std.error(st0_cpue,na.rm), StripedBass_SE =std.error(sb0_cpue,na.rm), TulePerch_SE =std.error(tp0_cpue,na.rm)) %>% 
  write.csv("~/GitHub/suisunmarsh/Ch1/Data_Ch1_Annual_Abun_Mean&SE_Age0.csv", row.names=FALSE)
```












###QUESTION 2. 
## 

## PEAK ABUNDANCE 
```{r, Part A. Means of means}
head(chapter1_biom2)

# mean abundance by year and month
peak_abundance <- chapter1_biom2 %>% 
  select(Yr, Mo, ST0_CPUE, SB0_CPUE, TP0_CPUE) %>% 
  group_by(Yr, Mo) %>% 
  summarize(Splittail = mean(ST0_CPUE), StripedBass = mean(SB0_CPUE), TulePerch = mean(TP0_CPUE), Splittail_SE =std.error(ST0_CPUE,na.rm), StripedBass_SE =std.error(SB0_CPUE,na.rm), TulePerch_SE =std.error(TP0_CPUE,na.rm))

peak_abundance_gathered <- peak_abundance %>% 
  select(Yr, Mo, Splittail:TulePerch) %>% 
  gather(Species, CPUE, Splittail:TulePerch)
View(peak_abundance_gathered)

# try filtering out just one species
peak_abundance_st0 <- peak_abundance_gathered %>% 
  filter(Species=="Splittail")

# box plot - mean of means
peak_boxplot <- ggplot(peak_abundance_gathered, aes(x=Mo, y=CPUE, group=Mo)) +
  geom_boxplot(outlier.size = 0.5) + 
  scale_x_discrete(name= "Month", breaks=c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  theme_economist_white() +
  scale_color_economist()
peak_boxplot + facet_grid(.~Species, scales = "free")
```

```{r, means of all data points}
peak_abundance_all <- chapter1_biom2 %>% 
  select(Mo, ST0_CPUE, SB0_CPUE, TP0_CPUE) %>% 
  rename(Splittail = ST0_CPUE, StripedBass = SB0_CPUE, TulePerch=TP0_CPUE ) %>% 
  gather(Species, CPUE, Splittail:TulePerch)
View(peak_abundance_all)

# box plot - mean of all data points
peak_boxplot <- ggplot(peak_abundance_all, aes(x=Mo, y=CPUE, group=Mo)) +
  geom_boxplot(outlier.size = 0.5) + 
  scale_x_discrete(name= "Month", breaks=c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  theme_economist_white() +
  scale_color_economist()
peak_boxplot + facet_grid(.~Species, scales = "free")

```



## PEAK BIOMASS - Part A
```{r}
biomass_gpue <- chapter1_biom2 %>% 
  select(YrMoSite, Yr, Mo, st0_b_tot, sb0_b_tot, tp0_b_tot, TowMin) %>% 
  group_by(YrMoSite) %>% 
  mutate(st0_gpue = st0_b_tot/ TowMin, sb0_gpue = sb0_b_tot/ TowMin, tp0_gpue = tp0_b_tot/ TowMin)

biomass_gpue2 <- biomass_gpue %>% 
  select(YrMoSite, st0_gpue, sb0_gpue, tp0_gpue)

# add GPUE values to dataset and write new file
chapter1_biom2_joined <- chapter1_biom2 %>% 
  inner_join(biomass_gpue2, by="YrMoSite")

chapter1_biom2_joined %>% 
  write_csv("~/GitHub/suisunmarsh/Ch1/Data_Ch1_Abun&Biomass.csv")

# calculate mean biomass by month
peak_biomass <- biomass_gpue %>% 
  select(Yr, Mo, st0_gpue, sb0_gpue, tp0_gpue) %>% 
  group_by(Yr, Mo) %>% 
  summarize(Splittail = mean(st0_gpue), StripedBass = mean(sb0_gpue), TulePerch = mean(tp0_gpue), Splittail_SE =std.error(st0_gpue,na.rm), StripedBass_SE =std.error(sb0_gpue,na.rm), TulePerch_SE =std.error(tp0_gpue,na.rm))

peak_biomass_gathered <- peak_biomass %>% 
  select(Yr, Mo, Splittail:TulePerch) %>% 
  gather(Species, GPUE, Splittail:TulePerch)
View(peak_biomass_gathered)



# box plot - mean of means
peak_boxplot2 <- ggplot(peak_biomass_gathered, aes(x=Mo, y=GPUE, group=Mo)) +
  geom_boxplot(outlier.size = 0.5) + 
  scale_x_discrete(name= "Month", breaks=c(1,2,3,4,5,6,7,8,9,10,11,12)) +
  theme_economist_white() +
  scale_color_economist()
peak_boxplot2 + facet_grid(.~Species, scales = "free")
```



