---
title: "Ch1_Abundance"
author: "Denise Colombano"
date: "February 8, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(ggplot2)
library(magrittr)
library(ggthemes)
library(gcookbook)
library(grid)
library(plotrix)
library(readr)

# import dataset that was created in the file Ch1_Biomass_Estimates, aka 'chapter1_biom_redo'
chapter1 <- read_csv("~/GitHub/suisunmarsh/Ch1/Data_output/Data_Ch1_BiomassEstimates_1995-2016_Redo.csv")

View(chapter1)
glimpse(chapter1)
chapter1$id <- 1:5475
str(chapter1)
summary(chapter1)
```

```{r, catch summaries of age-0}
yoy <- chapter1 %>% 
  summarize(sum(`ST_Age-0`),sum(`SB_Age-0`),sum(`TP_Age-0`))

st_compare <- chapter1 %>% 
  summarize(sum(`ST_Age-0`), sum(`ST_Age-1`), sum(`ST_Age-2+`))

sb_compare <- chapter1 %>% 
  summarize(sum(`SB_Age-0`), sum(`SB_Age-1`), sum(`SB_Age-2+`))

sf_compare <- chapter1 %>% 
  summarize(sum(`SF_Age-0`), sum(`SF_Age-1`), sum(`SF_Age-2+`))

tp_compare <- chapter1 %>% 
  summarize(sum(`TP_Age-0`), sum(`TP_Age-1+`))
```

## Create DF and CSV file for months May through September 1995-2016.
```{r, create data frames}
# create dataset with variables of interest

#Use `select` & `filter`:

 #- can select columns by names, attributes
 #- filter by rows, can filter by values

# Make a dataset using pipes (ctrl + shift + M) in magrittr

# filter out months 1-4 and 10-12
chapter1<- chapter1 %>%
  filter(Mo>4) #%>% 
  #filter(Mo<10)

summary(chapter1)

# create new dataframe for YOY CPUE
chapter1_yoy <- chapter1 %>% 
  select(Slough, YearMo, Mo, Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  group_by(Slough, Yr) %>% 
  summarize(st0_cpue = sum(`ST_Age-0`/ TowMin), sb0_cpue = sum(`SB_Age-0`/ TowMin), tp0_cpue = sum(`TP_Age-0`/ TowMin)) #%>% 
  #write_csv("~/GitHub/suisunmarsh/Ch1/Data_output/Age0_CPUE_1995-2016_Redo.csv")

# this is the NEW version "redo" with updated DV2 and SU1 data
abundance <- read_csv("~/GitHub/suisunmarsh/Ch1/Data_output/Age0_CPUE_1995-2016_Redo.csv")
glimpse(abundance)
View(abundance)
```

```{r, BIOMASS}

```

## DIFFERENCE BAR CHART
## Plot annual YOY abundance using relative differences and means
```{r, mean abundance}
## Calculate mean abundance by year
# Group by and Summarize
mean_abundance <- abundance %>%
  group_by(Yr) %>% 
  summarize(st0_mean = mean(st0_cpue), sb0_mean = mean(sb0_cpue), tp0_mean = mean(tp0_cpue))
View(mean_abundance)
```

```{r, join back with chapter1 dataframe}

# turn zero CPUE into NA so doesn't show up as negative values on chart
abundance_nozeros <- abundance %>% mutate(st0_cpue = "is.na<-"(st0_cpue, st0_cpue == 0)) %>% 
  mutate(sb0_cpue = "is.na<-"(sb0_cpue, sb0_cpue == 0)) %>% 
  mutate(tp0_cpue = "is.na<-"(tp0_cpue, tp0_cpue == 0))

# inner join
abundance_joined <- abundance_nozeros %>%
  inner_join(mean_abundance, by="Yr")
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
abundance_differences <- abundance_joined %>% 
  mutate(Splittail = st0_cpue - st0_mean, `Striped Bass` = sb0_cpue - sb0_mean, `Tule Perch` = tp0_cpue - tp0_mean) %>%
  select(Slough, Yr, Splittail, `Striped Bass`, `Tule Perch`) %>% 
  glimpse

a <- abundance_differences # shorthand
View(a)
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
a$Slough <- as.factor(a$Slough)
a$Yr <- as.factor(a$Yr)
#a$YrSlough <- as.factor(a$YrSlough)

# tidyr gather
a_gathered <- a %>% 
  gather(Species, Relative_CPUE, Splittail:`Tule Perch`)
View(a_gathered)

# set levels for species (sb, st, tp)
a_gathered$Species=factor(a_gathered$Species,levels=c("Striped Bass", "Splittail", "Tule Perch"))
levels(a_gathered$Species)
```

Plot relative differences:
First let's look at disproportionately greater than average cPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, mean difference bar charts}
# with legend
colors1 <- c("grey50","grey20")

a_gathered$Above_Annual_Mean <- a_gathered$Relative_CPUE >= 0
bar <- ggplot(a_gathered, aes(x=Yr, y=Relative_CPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity", colour="black", size=0.2) +
  theme_bw() +
  scale_fill_manual(values = colors1)+
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Catch Per Minute (Annual Mean=0)", x="", title="Relative Annual Catch of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 1), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=12), axis.title.x = element_text(size=12), axis.title.y = element_text(size=12), title = element_text(size=10)) # center title

# without legend
a_gathered$Above_Annual_Mean <- a_gathered$Relative_CPUE >= 0
bar <- ggplot(a_gathered, aes(x=Yr, y=Relative_CPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity", colour="black", size=0.2) +
  theme_bw() +
  scale_fill_manual(values = colors1)+
  guides(fill=FALSE) +
  scale_x_discrete(name= "", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Catch Per Minute (Adjusted Annual Mean=0)", x="")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm"), axis.title.x = element_text(size=18), axis.title.y = element_text(size=12), axis.line.y = element_line(size=0.5), strip.text = element_text(size=12), strip.background = element_rect(colour="black", fill="white", size=0.5))
        
#panel.grid.major = element_blank()

```


Then let's compare differences in pos/neg to relative proportion CPUE:
Notice how the "disproportionate" differences aren't obvious in this plot because
It's not standardized by annual averages.
```{r, bar chart as % total annual catch}
# summarize total annual catches
abundance_annual <- abundance %>% 
  select(Slough, Yr, st0_cpue, sb0_cpue, tp0_cpue) %>% 
  group_by(Yr) %>% 
  summarize(st0_annual = sum(st0_cpue), sb0_annual = sum(sb0_cpue), tp0_annual = sum(tp0_cpue)) %>% 
  glimpse

# inner join
abundance_totals <- abundance %>%
  inner_join(abundance_annual, by="Yr")
View(abundance_totals)

# create new variables (relative catch)
abundance_totals$st0_perc <- (abundance_totals$st0_cpue/ abundance_totals$st0_annual)
abundance_totals$sb0_perc <- (abundance_totals$sb0_cpue/ abundance_totals$sb0_annual)
abundance_totals$tp0_perc <- (abundance_totals$tp0_cpue/ abundance_totals$tp0_annual)

# Convert to factors first
abundance_totals$Slough <- as.factor(abundance_totals$Slough)
abundance_totals$Yr <- as.factor(abundance_totals$Yr)
abundance_totals$YrSlough <- as.factor(abundance_totals$YrSlough)

# tidyr gather
abundance_totals_renamed <- rename(abundance_totals, Splittail=st0_cpue, StripedBass=sb0_cpue, TulePerch=tp0_cpue)

abundance_totals_gathered <- abundance_totals_renamed %>% 
  select(Slough, Yr, Splittail, StripedBass, TulePerch) %>% 
  gather(Species, CPUE, Splittail:TulePerch)
View(abundance_totals_gathered)

# add habitat - from code further down
abundance_totals_habitat <- abundance_totals_gathered %>% 
  left_join(habitat, by="Slough")
```

```{r, mean YOY CPUE box plots}
# box plot
box <- ggplot(abundance_totals_gathered, aes(x=Slough, y=CPUE)) + 
  geom_boxplot(outlier.size = 0.5) +
  labs(y="Catch Per Minute", x="", title="Young-of-Year Catch by Slough") +
  theme_economist_white() +
  scale_fill_economist() +
  guides(fill=FALSE)

box + facet_grid(Species~., scales="free") + theme(plot.title = element_text(hjust = 0.5), title = element_text(size=18), axis.text.x = element_text(size=18), axis.title.y = element_text(size=18), axis.text.y = element_text(size=18) )
```


## Plot Mean YOY CPUE
```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Tidy data

# Convert to factors first
abundance$Slough <- as.factor(abundance$Slough)
abundance$Yr <- as.factor(abundance$Yr)
abundance$YrSlough <- as.factor(abundance$YrSlough)

# tidyr gather
abundance_gathered <- abundance %>% 
  gather(Species, Mean_CPUE, st0_cpue:tp0_cpue) 
View(abundance_gathered)
```


```{r, bar charts for mean YOY cpue}
bar <- ggplot(abundance_gathered, aes(x = Yr, y = Mean_CPUE)) +
  geom_bar(stat="identity") +
  scale_colour_economist() +
  theme_economist_white()
bar + theme(axis.text.x = element_text(angle = 90, hjust = 1)) + facet_grid(Species~Slough, scales="free") 
```

```{r, mean YOY CPUE stacked bar chart}
## Catch
# Group by and Summarize
mean_abundance_slough <- abundance %>%
  group_by(Slough) %>% 
  summarize(Splittail = mean(st0_cpue), StripedBass = mean(sb0_cpue), TulePerch = mean(tp0_cpue))
View(mean_abundance_slough)

# tidyr gather
slough_gathered <- mean_abundance_slough %>% 
  gather(Species, Mean_CPUE, Splittail:TulePerch) 
View(slough_gathered)

# stacked bar 1: abundance
stacked_bar <- ggplot(slough_gathered, aes(x=reorder(Slough, Mean_CPUE), y=Mean_CPUE, fill=Species)) +
  geom_bar(stat="identity", size=0.25) +
  scale_fill_economist() +
  theme_economist_white() +
  coord_flip() +
  guides(fill=guide_legend(reverse=TRUE)) +
  labs(y="Average Catch Per Minute", x="", title="Young-of-Year Catch by Slough")
stacked_bar + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=12), axis.title.x = element_text(size=12), title = element_text(size=10)) # center title
```


## Plot total YOY CPUE
```{r, total YOY CPUE stacked bar charts}
## Catch
# Group by and Summarize
sum_abundance_slough <- chapter1 %>%
  select(Slough, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  group_by(Slough) %>% 
  summarize(Splittail = (sum(`ST_Age-0`)/ sum(TowMin)), StripedBass = (sum(`SB_Age-0`)/ sum(TowMin)), TulePerch = (sum(`TP_Age-0`)/ sum(TowMin)))
View(sum_abundance_slough)

# tidyr gather
sum_slough_gathered <- sum_abundance_slough %>% 
  gather(Species, Total_CPUE, Splittail:TulePerch) 
View(sum_slough_gathered)

# stacked bar 3: total abundance
stacked_bar3 <- ggplot(sum_slough_gathered, aes(x=reorder(Slough, -Total_CPUE), y=Total_CPUE, fill=Species)) +
  geom_bar(stat="identity", size=0.25) +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  labs(y="Catch Per Minute", x="", title="Young-of-Year by Slough")
stacked_bar3 + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=12), axis.title.x = element_text(size=12), axis.title.y = element_text(size=12),title = element_text(size=10)) # center title
```

## All Age Classes
```{r, age class abundances}
ageclass_abundance <- chapter1 %>% 
  select(Slough, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, `ST_Age-1`, `SB_Age-1`, `TP_Age-1+`, `ST_Age-2+`, `SB_Age-2+`, TowMin) %>% 
  group_by(Slough) %>% 
  summarize(ST_YOY=sum(`ST_Age-0`)/ sum(TowMin), SB_YOY=sum(`SB_Age-0`)/ sum(TowMin), TP_YOY=sum(`TP_Age-0`)/ sum(TowMin), ST_Older = sum(`ST_Age-1`+ `ST_Age-2+`)/ sum(TowMin), SB_Older=sum(`SB_Age-1`+ `SB_Age-2+`)/ sum(TowMin), TP_Older=sum(`TP_Age-1+`)/ sum(TowMin)) 
glimpse(ageclass_abundance)
summary(ageclass_abundance)

# tidyr
ageclass_abundance$Slough <- as.factor(ageclass_abundance$Slough)

# tidyr gather
ageclass_abundance_gathered <- ageclass_abundance %>% 
  select(Slough, ST_YOY:TP_Older) %>% 
  gather(`Species+Stage`, CPUE, ST_YOY:TP_Older)
View(ageclass_abundance_gathered)

stacked_bar4 <- ggplot(ageclass_abundance_gathered, aes(x=reorder(Slough, -CPUE), y=CPUE, fill=`Species+Stage`)) +
  geom_bar(stat="identity", size=0.25) +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  labs(y="Catch Per Minute", x="", title="Total Catch by Slough")
stacked_bar4 + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title
```



## Stacked Area Plots
# All Ages CPUE
```{r, stacked area graph by Yr}
ageclass_abundance_yr <- chapter1 %>% 
  select(Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, `ST_Age-1`, `SB_Age-1`, `TP_Age-1+`, `ST_Age-2+`, `SB_Age-2+`, TowMin) %>% 
  group_by(Yr) %>% 
  summarize(`Striped Bass YOY`=sum(`SB_Age-0`)/ sum(TowMin), `Splittail YOY`=sum(`ST_Age-0`)/ sum(TowMin), `Tule Perch YOY`=sum(`TP_Age-0`)/ sum(TowMin), `Striped Bass Older`=sum(`SB_Age-1`+ `SB_Age-2+`)/ sum(TowMin), `Splittail Older` = sum(`ST_Age-1`+ `ST_Age-2+`)/ sum(TowMin), `Tule Perch Older`=sum(`TP_Age-1+`)/ sum(TowMin)) 
glimpse(ageclass_abundance_yr)
summary(ageclass_abundance_yr)

# tidyr
ageclass_abundance_yr$Yr <- as.factor(ageclass_abundance_yr$Yr)

# tidyr gather
ageclass_abundance_gathered_yr <- ageclass_abundance_yr %>% 
  select(Yr, `Striped Bass YOY`:`Tule Perch Older`) %>% 
  gather(`Species+Stage`, CPUE, `Striped Bass YOY`:`Tule Perch Older`)
View(ageclass_abundance_gathered_yr)
glimpse(ageclass_abundance_gathered_yr)

# change levels of Species+Stage factor for plotting control
ageclass_abundance_gathered_yr$`Species+Stage`=factor(ageclass_abundance_gathered_yr$`Species+Stage`,levels=c("Striped Bass Older", "Striped Bass YOY", "Splittail Older", "Splittail YOY", "Tule Perch Older", "Tule Perch YOY"))
levels(ageclass_abundance_gathered_yr$`Species+Stage`)

# stacked area plot all species
colors <- c("black","grey20","grey50","grey70", "grey90", "white")

stacked_area <- ggplot(ageclass_abundance_gathered_yr, aes(x=Yr, y=CPUE, fill=`Species+Stage`, group=`Species+Stage`)) +
  geom_area(stat="identity", colour="black", size=0.2) +
  guides(fill=guide_legend(reverse=FALSE)) +
  scale_fill_manual(values=colors)+
  theme_bw()+
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Total Catch of Species & Life Stage by Year")
stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title


?scale_fill_manual
```

## SPLITTAIL
```{r, stacked area plots for SPLITTAIL}
## SPLITTAIL

## YOY vs. OLDER
ageclass_abundance_gathered_st <- ageclass_abundance_yr %>% 
  select(Yr, ST_YOY, ST_Older) %>% 
  gather(`Species+Stage`, CPUE, ST_YOY, ST_Older)

# stacked area plot
stacked_area <- ggplot(ageclass_abundance_gathered_st, aes(x=Yr, y=CPUE, fill=`Species+Stage`, group=`Species+Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Total Catch of Species & Life Stage by Year")
stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title

## Age-0 VS. Age-1 VS. Age-2+
ageclass_abundance_st <- chapter1 %>% 
  select(Yr, `ST_Age-0`, `ST_Age-1`, `ST_Age-2+`, TowMin) %>% 
  group_by(Yr) %>% 
  summarize(`Age-0`=sum(`ST_Age-0`)/sum(TowMin), `Age-1` = sum(`ST_Age-1`)/ sum(TowMin),`Age-2+`=sum(`ST_Age-2+`)/ sum(TowMin) )

# tidyr
ageclass_abundance_st$Yr <- as.factor(ageclass_abundance_st$Yr)

# tidyr gather
ageclass_abundance_gathered_st <- ageclass_abundance_st %>% 
  select(Yr, `Age-0`, `Age-1`, `Age-2+`) %>% 
  gather(`Splittail+Stage`, CPUE, `Age-0`:`Age-2+`)

ageclass_abundance_gathered_st$`Stage` <- factor(ageclass_abundance_gathered_st$`Splittail+Stage`, levels=c("Age-2+", "Age-1", "Age-0"))

stacked_area <- ggplot(ageclass_abundance_gathered_st, aes(x=Yr, y=CPUE, fill=`Stage`, group=`Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Sacramento Splittail CPUE")

stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title

```

## STRIPED BASS
```{r, stacked area plots for STRIPED BASS}
## STRIPED BASS

## YOY vs. OLDER
ageclass_abundance_gathered_sb <- ageclass_abundance_yr %>% 
  select(Yr, SB_YOY, SB_Older) %>% 
  gather(`StripedBass+Stage`, CPUE, SB_YOY, SB_Older)

# stacked area plot
stacked_area <- ggplot(ageclass_abundance_gathered_sb, aes(x=Yr, y=CPUE, fill=`StripedBass+Stage`, group=`StripedBass+Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Striped Bass CPUE")
stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title


## Age-0 VS. Age-1 VS. Age-2+
ageclass_abundance_sb <- chapter1 %>% 
  select(Yr, `SB_Age-0`, `SB_Age-1`, `SB_Age-2+`, TowMin) %>% 
  group_by(Yr) %>% 
  summarize(`Age-0`=sum(`SB_Age-0`)/sum(TowMin), `Age-1` = sum(`SB_Age-1`)/ sum(TowMin),`Age-2+`=sum(`SB_Age-2+`)/ sum(TowMin) )

# tidyr
ageclass_abundance_sb$Yr <- as.factor(ageclass_abundance_sb$Yr)

# tidyr gather
ageclass_abundance_gathered_sb <- ageclass_abundance_sb %>% 
  select(Yr, `Age-0`, `Age-1`, `Age-2+`) %>% 
  gather(`StripedBass+Stage`, CPUE, `Age-0`:`Age-2+`)

ageclass_abundance_gathered_sb$`Stage` <- factor(ageclass_abundance_gathered_sb$`StripedBass+Stage`, levels=c("Age-2+", "Age-1", "Age-0"))

stacked_area <- ggplot(ageclass_abundance_gathered_sb, aes(x=Yr, y=CPUE, fill=`Stage`, group=`Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Striped Bass CPUE")

stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title
```

## TULE PERCH
```{r, stacked area plots for TULE PERCH}
## TULE PERCH

## YOY vs. OLDER
ageclass_abundance_gathered_tp <- ageclass_abundance_yr %>% 
  select(Yr, TP_YOY, TP_Older) %>% 
  gather(`TulePerch+Stage`, CPUE, TP_YOY, TP_Older)

# stacked area plot
stacked_area <- ggplot(ageclass_abundance_gathered_sb, aes(x=Yr, y=CPUE, fill=`TulePerch+Stage`, group=`TulePerch+Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Total Catch of Species & Life Stage by Year")
stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title


## Age-0 VS. Age-1+
ageclass_abundance_tp <- chapter1 %>% 
  select(Yr, `TP_Age-0`, `TP_Age-1+`, TowMin) %>% 
  group_by(Yr) %>% 
  summarize(`Age-0`=sum(`TP_Age-0`)/sum(TowMin), `Age-1+` = sum(`TP_Age-1+`)/ sum(TowMin) )

# tidyr
ageclass_abundance_tp$Yr <- as.factor(ageclass_abundance_tp$Yr)

# tidyr gather
ageclass_abundance_gathered_tp <- ageclass_abundance_tp %>% 
  select(Yr, `Age-0`, `Age-1+`) %>% 
  gather(`TulePerch+Stage`, CPUE, `Age-0`:`Age-1+`)

ageclass_abundance_gathered_tp$`Stage` <- factor(ageclass_abundance_gathered_tp$`TulePerch+Stage`, levels=c( "Age-1+", "Age-0"))

stacked_area <- ggplot(ageclass_abundance_gathered_tp, aes(x=Yr, y=CPUE, fill=`Stage`, group=`Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Tule Perch CPUE")

stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title
```

## YOY Only
```{r}
yoy_abundance_yr <- chapter1 %>% 
  select(Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  group_by(Yr) %>% 
  summarize(Splittail=sum(`ST_Age-0`)/ sum(TowMin), StripedBass=sum(`SB_Age-0`)/ sum(TowMin), TulePerch=sum(`TP_Age-0`)/ sum(TowMin))
glimpse(yoy_abundance_yr)
summary(yoy_abundance_yr)

# tidyr
yoy_abundance_yr$Yr <- as.factor(yoy_abundance_yr$Yr)

# tidyr gather
yoy_abundance_gathered_yr <- yoy_abundance_yr %>% 
  select(Yr, Splittail:TulePerch) %>% 
  gather(Species, CPUE, Splittail:TulePerch)
View(yoy_abundance_gathered_yr)
glimpse(yoy_abundance_gathered_yr)

# stacked area plot
stacked_area <- ggplot(yoy_abundance_gathered_yr, aes(x=Yr, y=CPUE, fill=Species, group=Species)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Total Catch of Young-of-Year")
stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title
```


## Abundance by habitat
```{r, habitat characteristics}
## Habitat
habitat <- read_csv("~/GitHub/suisunmarsh/data/Suisun_Marsh_Habitat.csv")
View(habitat)
summary(habitat)

# Convert to factors first
habitat$Slough <- as.factor(habitat$Slough)
habitat$Habitat <- as.factor(habitat$Habitat)

# Rename column names for plots
habitat$TidalMarsh <- habitat$tidal_perc
habitat$Channel <- habitat$channel_perc
habitat$ManagedMarsh <- habitat$managed_perc
habitat$Upland <- habitat$upland_perc

#habitats <- habitat %>%
  #select(Slough, tidal, managed, upland, land) %>% 
  #group_by(Slough) %>% 
  #summarize(Tidal = tidal/land, Managed = managed/land, Upland = upland/land)

habitat_gathered <- habitat %>% 
  select(Habitat, Slough, TidalMarsh:Upland) %>% 
  gather(Habitats, Proportion, TidalMarsh:Upland)
View(habitat_gathered)

habitat_gathered$Habitats <- as.factor(habitat_gathered$Habitats)

# assign levels for sloughs in order to reorder them in the x-axis
habitat_gathered$levels <- factor(habitat_gathered$Habitats)
habitat_gathered$levels <- factor(habitat_gathered$Habitats, levels=c("TidalMarsh", "ManagedMarsh", "Upland", "Channel"))
habitat_gathered$levels <- factor(habitat_gathered$levels, levels=c("Montezuma", "Boynton", "Cutoff", "Nurse", "Peytonia", "SuisunLower", "SuisunUpper", "Goodyear", "Spring Branch", "Denverton"))
```

```{r, box plot for habitat attributes}
box2 <- ggplot(habitat_gathered, aes(x=Habitats, y=Proportion, fill=Habitats)) + 
  geom_boxplot(outlier.size = 0.5) +
  labs(y="Percent Land Cover (200m Buffer)", x="", title="Land Cover Types") +
  theme_economist_white() +
  scale_fill_economist() +
  guides(fill=FALSE)
box2 + facet_grid(Habitat~.) + theme(plot.title = element_text(hjust = 0.5), title = element_text(size=10), axis.text.x = element_text(size=12), axis.title.y = element_text(size=12) )
```

### Season 1
## Bar chart of YOY by slough and Season with error bars
```{r}
??plotrix

# calculate mean CPUE and SE per slough
# use plotrix pacakge to calculate SE

# SUMMER Months 5,6,7,8
chapter1_yoy_season <- chapter1 %>%
  select(Slough, YearMo, Mo, Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  mutate(`ST0_CPUE`=`ST_Age-0`/TowMin, `SB0_CPUE`=`SB_Age-0`/TowMin, `TP0_CPUE`=`TP_Age-0`/TowMin)

chapter1_yoy_season <- mutate(chapter1_yoy_season, Season = ifelse(Mo > 8, "Autumn", "Summer"))

slough_cpue <- chapter1_yoy_season %>% 
  select(Slough, Season, `ST0_CPUE`, `SB0_CPUE`, `TP0_CPUE`) %>% 
  group_by(Slough, Season) %>% 
  summarize(Splittail = mean(`ST0_CPUE`), `Striped Bass` = mean(`SB0_CPUE`), `Tule Perch` = mean(`TP0_CPUE`), Splittail_SE =std.error(`ST0_CPUE`,na.rm), StripedBass_SE =std.error(`SB0_CPUE`,na.rm), TulePerch_SE =std.error(`TP0_CPUE`,na.rm))

# convert these to long format using gather and join
slough_cpue_mean <- slough_cpue %>% 
  select(Slough,Season, Splittail:`Tule Perch`) %>% 
  gather(Species, CPUE_Mean, Splittail:`Tule Perch`) 
View(slough_cpue_mean)

slough_cpue_mean_se <- slough_cpue %>% 
  select(Slough, Season,Splittail_SE:TulePerch_SE) %>% 
  rename(Splittail=Splittail_SE, `Striped Bass`=StripedBass_SE, `Tule Perch`=TulePerch_SE) %>% 
  gather(Species, CPUE_SE, Splittail:`Tule Perch`)  %>% 
  inner_join(slough_cpue_mean, by=c("Slough", "Species", "Season"))
View(slough_cpue_mean_se)

slough_cpue_mean_se$Species=factor(slough_cpue_mean_se$Species,levels=c("Striped Bass", "Splittail", "Tule Perch"))
levels(slough_cpue_mean_se$Species)

slough_cpue_mean_se$Season=factor(slough_cpue_mean_se$Season,levels=c("Summer", "Autumn"))
levels(slough_cpue_mean_se$Season)

# plot a bar chart with error bars
bar2 <- ggplot(slough_cpue_mean_se, aes(x=Season, y=CPUE_Mean, fill=Slough)) +
  geom_bar(stat="identity", color="black")+
  theme_economist_white() +
  scale_fill_economist()+
  labs(y="Mean Catch Per Minute", x="", title="Mean YOY CPUE by Season")+
  scale_x_discrete(breaks=c("Summer", "Autumn"), labels=c("Summer", "Autumn"))+
  guides(fill=FALSE)
bar2 + geom_errorbar(aes(ymin = CPUE_Mean - CPUE_SE, ymax = CPUE_Mean + CPUE_SE, width = 0.5)) + facet_grid(Species~Slough, scales = "free")+ theme(plot.title = element_text(hjust = 0.5), axis.title.y = element_text(size=18),title = element_text(size=18)) 
```

### Season 2
## Horizontal Bar Difference chart of YOY by slough and Season with error bars
```{r}
# turn zero CPUE into NA so doesn't show up as negative values on chart
chapter1_yoy_season_nozeros <- chapter1_yoy_season %>% 
  mutate(ST0_CPUE = "is.na<-"(ST0_CPUE , ST0_CPUE == 0)) %>% 
  mutate(SB0_CPUE = "is.na<-"(SB0_CPUE, SB0_CPUE == 0)) %>% 
  mutate(TP0_CPUE = "is.na<-"(TP0_CPUE, TP0_CPUE == 0)) %>% 
  select(Slough:Yr, ST0_CPUE:Season)

# mean season
chapter1_yoy_season_sum <- chapter1_yoy_season %>%
  select(Slough, Season, `ST_Age-0`:`TP_Age-0`, TowMin) %>% 
  group_by(Slough,Season) %>% 
  summarize(ST0_Sum=sum(`ST_Age-0`)/sum(TowMin), SB0_Sum=sum(`SB_Age-0`)/sum(TowMin), TP0_Sum=sum(`TP_Age-0`)/sum(TowMin))

# reshape data for all three species
season_spread <- chapter1_yoy_season_sum %>%
  select(Slough, Season, ST0_Sum) %>% 
  spread(Season, ST0_Sum) %>% 
  mutate(Species="Splittail")

season_spread2 <- chapter1_yoy_season_sum %>%
  select(Slough, Season, SB0_Sum) %>% 
  spread(Season, SB0_Sum) %>% 
  mutate(Species="Striped Bass")

season_spread3 <- chapter1_yoy_season_sum %>%
  select(Slough, Season, TP0_Sum) %>% 
  spread(Season, TP0_Sum) %>% 
  mutate(Species="Tule Perch")

season_joined <- bind_rows(season_spread, season_spread2, season_spread3)
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
seasonal_differences <- season_joined %>% 
  mutate(Change = Autumn-Summer, Summer_neg =-Summer) 

s <- seasonal_differences
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
s$Slough <- as.factor(s$Slough)
s$Species <- as.factor(s$Species)

# set levels for species (sb, st, tp)
s$Species=factor(s$Species,levels=c("Striped Bass", "Splittail", "Tule Perch"))
levels(s$Species)

# create two dataframes for stacked plot
s_summer <- s %>% 
  select(Slough,Summer_neg,Species,Change)

s_autumn <- s %>% 
  select(Slough,Autumn,Species,Change)

# reshape data again (maybe use this, maybe not)
s_gathered <- s %>% 
  select(Slough, Species, Change, Autumn, Summer) %>% 
  gather(Season, CPUE, Autumn:Summer)
```

Plot relative differences:
First let's look at disproportionately greater than average cPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, seasonal difference bar charts}
# without legend

bar4 <- ggplot()+
  geom_bar(data=s_summer, aes(x=reorder(Slough, -Summer_neg), y=Summer_neg, fill="lightblue"), stat="identity") +
  geom_bar(data=s_autumn, aes(x=reorder(Slough, -Autumn), y=Autumn, fill="steelblue"), stat="identity") +
  coord_flip()+
  theme_economist_white()+
  scale_fill_excel()+
  guides(fill=FALSE)+
  labs(title="YOY Catch by Season: Summer vs. Autumn", x="", y="Catch Per Minute")
bar4 + facet_grid(.~Species, scales="free") + theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=18), axis.title.y = element_text(size=18), title = element_text(size=14), axis.line.y = element_line(size=0.5), strip.text = element_text(size=16), axis.text.y = element_text(size=16), axis.text.x = element_text(size=16))
```



## Mean annual YOY abundance plots by slough
```{r, calculate average annual abundance over 21 years}
## Calculate average ANNUAL cPUE (21 data points)

# calculate mean CPUE and SE per slough
# use plotrix pacakge to calculate SE
annual_mean_cpue <- abundance %>% 
  group_by(Slough) %>% 
  summarize(Splittail = mean(st0_cpue), StripedBass = mean(sb0_cpue), TulePerch = mean(tp0_cpue), Splittail_SE =std.error(st0_cpue,na.rm), StripedBass_SE =std.error(sb0_cpue,na.rm), TulePerch_SE =std.error(tp0_cpue,na.rm)) #%>% 
  #write.csv("~/GitHub/suisunmarsh/Ch1/Data_output/Data_Ch1_Annual_Abun_Mean&SE_Age0.csv", row.names=FALSE)

# convert these to long format 
slough_cpue_mean2 <- annual_mean_cpue %>% 
  select(Slough, Splittail:TulePerch) %>% 
  gather(Species, CPUE_Mean, Splittail:TulePerch) 
View(slough_cpue_mean2)

slough_cpue_mean_se2 <- annual_mean_cpue %>% 
  select(Slough, Splittail_SE:TulePerch_SE) %>% 
  rename(Splittail=Splittail_SE, StripedBass=StripedBass_SE, TulePerch=TulePerch_SE) %>% # same name for join
  select(Slough, Splittail:TulePerch) %>% 
  gather(Species, CPUE_SE, Splittail:TulePerch) 

# join them back together
slough_cpue_joined <- slough_cpue_mean2 %>% 
  inner_join(slough_cpue_mean_se2, by=c("Slough", "Species"))
View(slough_cpue_joined)

# plot a bar chart with error bars
bar3 <- ggplot(slough_cpue_joined, aes(x=Slough, y=CPUE_Mean, fill=Slough)) +
  geom_bar(stat="identity", position="dodge", colour="black") +
  facet_grid(Species~., scales="free")+
  theme_economist_white()+
  scale_fill_economist(guide=FALSE) +
  labs(y="Catch Per Minute", x="", title="Young-of-Year CPUE")

bar3

bar3 + geom_errorbar(aes(ymin = CPUE_Mean - CPUE_SE, ymax = CPUE_Mean + CPUE_SE), size=0.5, width = 0.25, position=position_dodge(0.9)) + facet_grid(Species~., scales="free") + theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=18), title = element_text(size=16))
```

## Mean annual YOY abundance plots by SITE along the longitudinal axis of SB and DV
```{r, mean at site}
## Calculate average ANNUAL cPUE (21 data points)

abundance_yoy_site <- chapter1 %>%
  select(StationCode, Yr, TowMin, `SB_Age-0`, `ST_Age-0`, `TP_Age-0`) %>% 
  group_by(StationCode, Yr) %>% 
  summarize(sb0_cpue=sum(`SB_Age-0`)/sum(TowMin), st0_cpue=sum(`ST_Age-0`)/sum(TowMin), tp0_cpue=sum(`TP_Age-0`)/sum(TowMin))

abundance_yoy_site1 <- abundance_yoy_site %>% 
  filter(StationCode=="SB1") %>% 
  mutate(Slough="Spring Branch")

abundance_yoy_site2 <- abundance_yoy_site %>% 
  filter(StationCode=="SB2") %>% 
  mutate(Slough="Spring Branch")

abundance_yoy_site3 <- abundance_yoy_site %>% 
  filter(StationCode=="DV2") %>% 
  mutate(Slough="Denverton")

abundance_yoy_site4 <- abundance_yoy_site %>% 
  filter(StationCode=="DV3") %>% 
  mutate(Slough="Denverton")


abundance_yoy_site_bind <- bind_rows(abundance_yoy_site1, abundance_yoy_site2, abundance_yoy_site3, abundance_yoy_site4)

# calculate mean CPUE and SE per site
# use plotrix pacakge to calculate SE
site_mean_cpue <- abundance_yoy_site_bind %>% 
  group_by(Slough,StationCode) %>% 
  summarize(`Striped Bass` = mean(sb0_cpue), Splittail = mean(st0_cpue), `Tule Perch` = mean(tp0_cpue), StripedBass_SE =std.error(sb0_cpue,na.rm), Splittail_SE =std.error(st0_cpue,na.rm), TulePerch_SE =std.error(tp0_cpue,na.rm)) #%>% 
  #write.csv("~/GitHub/suisunmarsh/Ch1/Data_output/Data_Ch1_Annual_Abun_Mean&SE_Age0.csv", row.names=FALSE)

# convert these to long format 
site_cpue_mean2 <- site_mean_cpue %>% 
  select(Slough, StationCode, `Striped Bass`:`Tule Perch`) %>% 
  gather(Species, CPUE_Mean, `Striped Bass`:`Tule Perch`) 
View(site_cpue_mean2)

site_cpue_mean_se2 <- site_mean_cpue %>% 
  select(Slough, StationCode, StripedBass_SE:TulePerch_SE) %>% 
  rename(`Striped Bass`=StripedBass_SE, Splittail=Splittail_SE, `Tule Perch`=TulePerch_SE) %>% # same name for join
  select(StationCode, `Striped Bass`:`Tule Perch`) %>% 
  gather(Species, CPUE_SE, `Striped Bass`:`Tule Perch`) 

# join them back together
site_cpue_joined <- site_cpue_mean2 %>% 
  inner_join(site_cpue_mean_se2, by=c("Slough", "StationCode", "Species"))
View(site_cpue_joined)

# set levels for species (sb, st, tp)
site_cpue_joined$Species=factor(site_cpue_joined$Species,levels=c("Striped Bass", "Splittail", "Tule Perch"))
levels(site_cpue_joined$Species)

# plot a bar chart with error bars
bar4 <- ggplot(site_cpue_joined, aes(x=StationCode, y=CPUE_Mean, fill=StationCode)) +
  geom_bar(stat="identity", position="dodge", width=0.55, colour="black") +
  facet_grid(Species~Slough, scales="free")+
  theme_economist_white()+
  scale_fill_economist(guide=FALSE) +
  scale_x_discrete(breaks=c("DV2", "DV3", "SB1", "SB2"), labels = c("Upper (DV2)", "Lower (DV3)", "Upper (SB1)", "Lower (SB2)"))+
  labs(y="Catch Per Minute", x="", title="YOY Hotspots: Upper vs. Lower Slough")

bar4

bar4 + geom_errorbar(aes(ymin = CPUE_Mean - CPUE_SE, ymax = CPUE_Mean + CPUE_SE), size=0.5, width = 0.25, position=position_dodge(0.9)) + facet_grid(Species~Slough, scales="free") + theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=18), title = element_text(size=14))

```

```{r, boxplot of CPUE and Biomass by Months}

chapter1_mo <- chapter1 %>%
  select(YrMoSite, Mo, TowMin, `ST_Age-0`:`TP_Age-0`) %>% 
  group_by(YrMoSite) %>% 
  mutate(`Splittail YOY`=sum(`ST_Age-0`)/sum(TowMin), `Striped Bass YOY`=sum(`SB_Age-0`)/sum(TowMin), `Tule Perch YOY`=sum(`TP_Age-0`)/sum(TowMin))

chapter1_mo_gathered <- chapter1_mo %>%
  select(Mo,`Splittail YOY`:`Tule Perch YOY`) %>% 
  gather(`Species+Stage`, CPUE, `Splittail YOY`:`Tule Perch YOY`)

box <- ggplot(chapter1_mo_gathered, aes(x=Mo, y=CPUE, group=Mo)) +
  geom_boxplot()
box + facet_grid(`Species+Stage`~., scales="free")        
        
```

```{r, Test out the Mann Kendall Test}
library(Kendall)
library(xts)

abund_st <- abundance_annual %>% 
  select(Yr,st0_annual)
plot(abund_st)
acf(abund_st)

data(PrecipGL)
plot(PrecipGL)
lines(lowess(time(PrecipGL),PrecipGL),lwd=3, col=2)
acf(PrecipGL)

```

