---
title: "Ch1_Abundance"
author: "Denise Colombano"
date: "February 8, 2017"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(tidyverse)
library(ggplot2)
library(magrittr)
library(ggthemes)
library(gcookbook)
library(grid)
library(plotrix)

chapter1 <- read_csv("~/GitHub/suisunmarsh/Ch1/Data_Ch1_Abundance.txt")
View(chapter1)
glimpse(chapter1)
chapter1$id <- 1:5223
str(chapter1)
summary(chapter1)

```

## Create DF and CSV file for months May through December 1995-2015.
```{r, create data frames}
# create dataset with variables of interest

#Use `select` & `filter`:

 #- can select columns by names, attributes
 #- filter by rows, can filter by values

# Make a dataset using pipes (ctrl + shift + M) in magrittr

# filter out months 1-4
chapter1<- chapter1 %>%
  filter(Mo>4)

# add in biomass calculations

# create new dataframe
chapter1 %>% 
  select(Slough, YearMo, YrSlough, Mo, Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  filter(Mo > 4) %>% 
  group_by(YrSlough, Slough, Yr) %>% 
  summarize(st0_cpue = sum(`ST_Age-0`/ TowMin), sb0_cpue = sum(`SB_Age-0`/ TowMin), tp0_cpue = sum(`TP_Age-0`/ TowMin)) %>% 
  write.csv("~/GitHub/suisunmarsh/Ch1/abundance.csv", row.names=FALSE)

abundance <- read_csv("~/GitHub/suisunmarsh/Ch1/abundance.csv")
glimpse(abundance)
View(abundance)
```

## Plotting annual YOY abundance using relative differences and means
```{r, mean abundance}
## Calculate mean abundance by year
# Group by and Summarize
mean_abundance <- abundance %>%
  group_by(Yr) %>% 
  summarize(st0_mean = mean(st0_cpue), sb0_mean = mean(sb0_cpue), tp0_mean = mean(tp0_cpue))
View(mean_abundance)
```

```{r, join back with chapter1 dataframe}

# Following this lesson:

## **https://ismayc.github.io/moderndiver-book/5-manip.html#joining-data-frames**

# inner join
abundance_joined <- abundance %>%
  inner_join(mean_abundance, by="Yr")
```

```{r, create new columns for pos/neg differences}
## Calculate differences between CPUE and marsh-wide average (pos/neg)
abundance_differences <- abundance_joined %>% 
  mutate(Splittail = st0_cpue - st0_mean, Striped_Bass = sb0_cpue - sb0_mean, Tule_Perch = tp0_cpue - tp0_mean) %>%
  select(YrSlough, Slough, Yr, Splittail, Striped_Bass, Tule_Perch) %>% 
  glimpse

a <- abundance_differences # shorthand
View(a)
```

```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Convert to factors first
a$Slough <- as.factor(a$Slough)
a$Yr <- as.factor(a$Yr)
a$YrSlough <- as.factor(a$YrSlough)

# tidyr gather
a_gathered <- a %>% 
  gather(Species, Relative_CPUE, Splittail:Tule_Perch)
View(a_gathered)
```

Plot relative differences:
First let's look at disproportionately greater than average cPUE among Slough & Yr for each species,
Using the a_gathered dataframe, representing differences from the mean (0) for that year.

```{r, mean difference bar charts}
# with legend
a_gathered$Above_Annual_Mean <- a_gathered$Relative_CPUE >= 0
bar <- ggplot(a_gathered, aes(x=Yr, y=Relative_CPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity") +
  scale_colour_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Catch Per Minute (Annual Mean=0)", x="", title="Relative Annual Catch of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 1), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=12), axis.title.x = element_text(size=12), axis.title.y = element_text(size=12), title = element_text(size=10)) # center title

# without legend
a_gathered$Above_Annual_Mean <- a_gathered$Relative_CPUE >= 0
bar <- ggplot(a_gathered, aes(x=Yr, y=Relative_CPUE, fill=Above_Annual_Mean, guide=FALSE)) +
  geom_bar(stat="identity", position="identity") +
  theme_economist_white() +
  scale_fill_economist()+
  guides(fill=FALSE) +
  scale_x_discrete(name= "Year", breaks=c(1995, 2000, 2005, 2010, 2015))+
  labs(y="Catch Per Minute (Annual Mean=0)", x="", title="Relative Annual Catch of Young-of-Year")
bar + facet_grid(Species ~ Slough, scales="free") +
  theme(axis.text.x = element_text(angle = -90, hjust = 0.5), axis.ticks = element_line(size = 0.5), axis.ticks.length = unit(0.25, "cm")) +
  theme(plot.title = element_text(hjust = 0.5), axis.title.x = element_text(size=14), axis.title.y = element_text(size=12), title = element_text(size=10), axis.line.y = element_line(size=0.5), strip.text = element_text(size=10))
        
        #strip.background = element_rect(colour="black", size=0.5)) 
```


Then let's compare differences in pos/neg to relative proportion CPUE:
Notice how the "disproportionate" differences aren't obvious in this plot because
It's not standardized by annual averages.
```{r, bar chart as % total annual catch}
# summarize total annual catches
abundance_annual <- abundance %>% 
  select(Slough, Yr, st0_cpue, sb0_cpue, tp0_cpue) %>% 
  group_by(Yr) %>% 
  summarize(st0_annual = sum(st0_cpue), sb0_annual = sum(sb0_cpue), tp0_annual = sum(tp0_cpue)) %>% 
  glimpse

# inner join
abundance_totals <- abundance %>%
  inner_join(abundance_annual, by="Yr")
View(abundance_totals)

# create new variables (relative catch)
abundance_totals$st0_perc <- (abundance_totals$st0_cpue/ abundance_totals$st0_annual)
abundance_totals$sb0_perc <- (abundance_totals$sb0_cpue/ abundance_totals$sb0_annual)
abundance_totals$tp0_perc <- (abundance_totals$tp0_cpue/ abundance_totals$tp0_annual)

# Convert to factors first
abundance_totals$Slough <- as.factor(abundance_totals$Slough)
abundance_totals$Yr <- as.factor(abundance_totals$Yr)
abundance_totals$YrSlough <- as.factor(abundance_totals$YrSlough)

# tidyr gather
abundance_totals_renamed <- rename(abundance_totals, Splittail=st0_cpue, StripedBass=sb0_cpue, TulePerch=tp0_cpue)

abundance_totals_gathered <- abundance_totals_renamed %>% 
  select(Slough, Yr, Splittail, StripedBass, TulePerch) %>% 
  gather(Species, CPUE, Splittail:TulePerch)
View(abundance_totals_gathered)
```

```{r, mean YOY CPUE box plots}
# box plot
box <- ggplot(abundance_totals_gathered, aes(x=Slough, y=CPUE, fill=Slough)) + 
  geom_boxplot(outlier.size = 0.5) +
  labs(y="Catch Per Minute", x="", title="Young-of-Year Catch by Slough") +
  theme_economist_white() +
  scale_fill_economist() +
  guides(fill=FALSE)
box + facet_grid(Species~. , scales="free") + theme(plot.title = element_text(hjust = 0.5), title = element_text(size=12), axis.text.x = element_text(size=12), axis.title.y = element_text(size=14) )
```


## Plot Mean YOY CPUE
```{r, convert dataframe to from wide to long format}

# Following the R Cookbook:

## *http://www.cookbook-r.com/Manipulating_data/Converting_data_between_wide_and_long_format/**

# Tidy data

# Convert to factors first
abundance$Slough <- as.factor(abundance$Slough)
abundance$Yr <- as.factor(abundance$Yr)
abundance$YrSlough <- as.factor(abundance$YrSlough)

# tidyr gather
abundance_gathered <- abundance %>% 
  gather(Species, Mean_CPUE, st0_cpue:tp0_cpue) 
View(abundance_gathered)
```


```{r, bar charts for mean YOY cpue}
bar <- ggplot(abundance_gathered, aes(x = Yr, y = Mean_CPUE)) +
  geom_bar(stat="identity") +
  facet_grid(Species~Slough) +
  scale_colour_economist() +
  theme_economist_white()
bar + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

```{r, mean YOY CPUE stacked bar chart}
## Catch
# Group by and Summarize
mean_abundance_slough <- abundance %>%
  group_by(Slough) %>% 
  summarize(Splittail = mean(st0_cpue), StripedBass = mean(sb0_cpue), TulePerch = mean(tp0_cpue))
View(mean_abundance_slough)

# tidyr gather
slough_gathered <- mean_abundance_slough %>% 
  gather(Species, Mean_CPUE, Splittail:TulePerch) 
View(slough_gathered)

# stacked bar 1: abundance
stacked_bar <- ggplot(slough_gathered, aes(x=reorder(Slough, Mean_CPUE), y=Mean_CPUE, fill=Species)) +
  geom_bar(stat="identity", size=0.25) +
  scale_fill_economist() +
  theme_economist_white() +
  coord_flip() +
  guides(fill=guide_legend(reverse=TRUE)) +
  labs(y="Average Catch Per Minute", x="", title="Young-of-Year Catch by Slough")
stacked_bar + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=12), axis.title.x = element_text(size=12), title = element_text(size=10)) # center title
```


## Plot total YOY CPUE
```{r, total YOY CPUE stacked bar charts}
## Catch
# Group by and Summarize
sum_abundance_slough <- abundance %>%
  select(Slough, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, TowMin) %>% 
  group_by(Slough) %>% 
  summarize(Splittail = (sum(`ST_Age-0`)/ sum(TowMin)), StripedBass = (sum(`SB_Age-0`)/ sum(TowMin)), TulePerch = (sum(`TP_Age-0`)/ sum(TowMin)))
View(sum_abundance_slough)

# tidyr gather
sum_slough_gathered <- sum_abundance_slough %>% 
  gather(Species, Total_CPUE, Splittail:TulePerch) 
View(sum_slough_gathered)

# stacked bar 3: total abundance
stacked_bar3 <- ggplot(sum_slough_gathered, aes(x=reorder(Slough, -Total_CPUE), y=Total_CPUE, fill=Species)) +
  geom_bar(stat="identity", size=0.25) +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  labs(y="Catch Per Minute", x="", title="Young-of-Year by Slough")
stacked_bar3 + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=12), axis.title.x = element_text(size=12), axis.title.y = element_text(size=12),title = element_text(size=10)) # center title
```

## All Age Classes
```{r, age class abundances}
ageclass_abundance <- chapter1 %>% 
  select(Slough, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, `ST_Age-1`, `SB_Age-1`, `TP_Age-1+`, `ST_Age-2+`, `SB_Age-2+`, TowMin) %>% 
  group_by(Slough) %>% 
  summarize(ST_YOY=sum(`ST_Age-0`)/ sum(TowMin), SB_YOY=sum(`SB_Age-0`)/ sum(TowMin), TP_YOY=sum(`TP_Age-0`)/ sum(TowMin), ST_Older = sum(`ST_Age-1`+ `ST_Age-2+`)/ sum(TowMin), SB_Older=sum(`SB_Age-1`+ `SB_Age-2+`)/ sum(TowMin), TP_Older=sum(`TP_Age-1+`)/ sum(TowMin)) 
glimpse(ageclass_abundance)
summary(ageclass_abundance)

# tidyr
ageclass_abundance$Slough <- as.factor(ageclass_abundance$Slough)

# tidyr gather
ageclass_abundance_gathered <- ageclass_abundance %>% 
  select(Slough, ST_YOY:TP_Older) %>% 
  gather(`Species+Stage`, CPUE, ST_YOY:TP_Older)
View(ageclass_abundance_gathered)

stacked_bar4 <- ggplot(ageclass_abundance_gathered, aes(x=reorder(Slough, -CPUE), y=CPUE, fill=`Species+Stage`)) +
  geom_bar(stat="identity", size=0.25) +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  labs(y="Catch Per Minute", x="", title="Total Catch by Slough")
stacked_bar4 + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title
```

```{r, stacked area graph by Yr}
ageclass_abundance_yr <- chapter1 %>% 
  select(Yr, `ST_Age-0`, `SB_Age-0`, `TP_Age-0`, `ST_Age-1`, `SB_Age-1`, `TP_Age-1+`, `ST_Age-2+`, `SB_Age-2+`, TowMin) %>% 
  group_by(Yr) %>% 
  summarize(ST_YOY=sum(`ST_Age-0`)/ sum(TowMin), SB_YOY=sum(`SB_Age-0`)/ sum(TowMin), TP_YOY=sum(`TP_Age-0`)/ sum(TowMin), ST_Older = sum(`ST_Age-1`+ `ST_Age-2+`)/ sum(TowMin), SB_Older=sum(`SB_Age-1`+ `SB_Age-2+`)/ sum(TowMin), TP_Older=sum(`TP_Age-1+`)/ sum(TowMin)) 
glimpse(ageclass_abundance_yr)
summary(ageclass_abundance_yr)

# tidyr
ageclass_abundance_yr$Yr <- as.factor(ageclass_abundance_yr$Yr)

# tidyr gather
ageclass_abundance_gathered_yr <- ageclass_abundance_yr %>% 
  select(Yr, ST_YOY:TP_Older) %>% 
  gather(`Species+Stage`, CPUE, ST_YOY:TP_Older)
View(ageclass_abundance_gathered_yr)
glimpse(ageclass_abundance_gathered_yr)

# stacked area plot
stacked_area <- ggplot(ageclass_abundance_gathered_yr, aes(x=Yr, y=CPUE, fill=`Species+Stage`, group=`Species+Stage`)) +
  geom_area(stat="identity") +
  scale_fill_economist() +
  theme_economist_white() +
  guides(fill=guide_legend(reverse=TRUE)) +
  scale_x_discrete(breaks=c(1995, 2000, 2005, 2010, 2015)) +
  labs(y="Catch Per Minute", x="", title="Total Catch of Species & Life Stage by Year")
stacked_area + theme(plot.title = element_text(hjust = 0.5), legend.title = element_text(face="italic", size=14), axis.title.y = element_text(size=14),title = element_text(size=10)) # center title
```

## Abundance by habitat
```{r, habitat characteristics}
## Habitat
habitat <- read_csv("~/GitHub/suisunmarsh/data/Suisun_Marsh_Habitat.csv")
View(habitat)
summary(habitat)

# Convert to factors first
habitat$Slough <- as.factor(habitat$Slough)
habitat$Habitat <- as.factor(habitat$Habitat)

# Rename column names for plots
habitat$TidalMarsh <- habitat$tidal_perc
habitat$Channel <- habitat$channel_perc
habitat$ManagedMarsh <- habitat$managed_perc
habitat$Upland <- habitat$upland_perc

#habitats <- habitat %>%
  #select(Slough, tidal, managed, upland, land) %>% 
  #group_by(Slough) %>% 
  #summarize(Tidal = tidal/land, Managed = managed/land, Upland = upland/land)

habitat_gathered <- habitat %>% 
  select(Habitat, Slough, TidalMarsh:Upland) %>% 
  gather(Habitats, Proportion, TidalMarsh:Upland)
View(habitat_gathered)

habitat_gathered$Habitats <- as.factor(habitat_gathered$Habitats)

# assign levels for sloughs in order to reorder them in the x-axis
habitat_gathered$levels <- factor(habitat_gathered$Habitats)
habitat_gathered$levels <- factor(habitat_gathered$Habitats, levels=c("TidalMarsh", "ManagedMarsh", "Upland", "Channel"))
#habitat_gathered$levels <- factor(habitat_gathered$levels, levels=c("Montezuma", "Boynton", "Cutoff", "Nurse", "Peytonia", "SuisunLower", "SuisunUpper", "Goodyear", "Spring Branch", "Denverton"))
```

```{r, box plot for habitat attributes}
box2 <- ggplot(habitat_gathered, aes(x=Habitats, y=Proportion, fill=Habitats)) + 
  geom_boxplot(outlier.size = 0.5) +
  labs(y="Percent Land Cover (200m Buffer)", x="", title="Land Cover Types") +
  theme_economist_white() +
  scale_fill_economist() +
  guides(fill=FALSE)
box2 + facet_grid(Habitat~.) + theme(plot.title = element_text(hjust = 0.5), title = element_text(size=10), axis.text.x = element_text(size=12), axis.title.y = element_text(size=12) )
```

```{r}
??plotrix

# calculate mean CPUE and SE per slough
# use plotrix pacakge to calculate SE
slough_cpue <- chapter1 %>% 
  select(Slough, `ST0_CPUE`, `SB0_CPUE`, `TP0_CPUE`) %>% 
  group_by(Slough) %>% 
  summarize(Splittail = mean(`ST0_CPUE`), StripedBass = mean(`SB0_CPUE`), TulePerch = mean(`TP0_CPUE`), Splittail_SE =std.error(`ST0_CPUE`,na.rm), StripedBass_SE =std.error(`SB0_CPUE`,na.rm), TulePerch_SE =std.error(`TP0_CPUE`,na.rm))

# convert these to long format using gather and join
slough_cpue_mean <- slough_cpue %>% 
  select(Slough, Splittail:TulePerch) %>% 
  gather(Species, CPUE_Mean, Splittail:TulePerch) 
View(slough_cpue_mean)

slough_cpue_mean_se <- slough_cpue %>% 
  select(Slough, Splittail_SE:TulePerch_SE) %>% 
  gather(Species, CPUE_SE, Splittail_SE:TulePerch_SE) %>% 
  inner_join(slough_cpue_mean, by="Slough")
View(slough_cpue_mean_se)

# plot a bar chart with error bars
bar2 <- ggplot(slough_cpue_mean_se, aes(x=Slough, y=CPUE_Mean, fill=Slough)) +
  geom_bar(stat="identity")+
  theme_economist_white() +
  scale_fill_economist()
bar2 + geom_errorbar(aes(ymin = CPUE_Mean - CPUE_SE, ymax = CPUE_Mean + CPUE_SE, width = 0.5)) + facet_grid(Species.y~., scales = "free") 
```


## Mean annual YOY abundance plots by slough
```{r, calculate average annual abundance over 21 years}
## Calculate average ANNUAL cPUE (21 data points)

# calculate mean CPUE and SE per slough
# use plotrix pacakge to calculate SE
annual_mean_cpue <- abundance %>% 
  group_by(Slough) %>% 
  summarize(Splittail = mean(st0_cpue), StripedBass = mean(sb0_cpue), TulePerch = mean(tp0_cpue), Splittail_SE =std.error(st0_cpue,na.rm), StripedBass_SE =std.error(sb0_cpue,na.rm), TulePerch_SE =std.error(tp0_cpue,na.rm)) %>% 
  write.csv("~/GitHub/suisunmarsh/Ch1/Data_Ch1_Annual_Abun_Mean&SE_Age0.csv", row.names=FALSE)


# convert these to long format using gather and join
slough_cpue_mean2 <- annual_mean_cpue %>% 
  select(Slough, Splittail:TulePerch) %>% 
  gather(Species, CPUE_Mean, Splittail:TulePerch) 
View(slough_cpue_mean2)

slough_cpue_mean_se2 <- annual_mean_cpue %>% 
  select(Slough, Splittail_SE:TulePerch_SE) %>% 
  gather(Species, CPUE_SE, Splittail_SE:TulePerch_SE) %>% 
  inner_join(slough_cpue_mean2, by="Slough")
View(slough_cpue_mean_se2)

# plot a bar chart with error bars
bar3 <- ggplot(slough_cpue_mean_se2, aes(x=Slough, y=CPUE_Mean, fill=Slough)) +
  geom_bar(stat="identity")+
  theme_economist_white() +
  scale_fill_economist()
bar3 + geom_errorbar(aes(ymin = CPUE_Mean - CPUE_SE, ymax = CPUE_Mean + CPUE_SE, width = 0.5)) + facet_grid(Species.y~., scales = "free") 
```

